---
title: "Analysis of Longevity from Muscle dTORC1 Effector Flies"
author: ' Dave Bridges, Isabelle Hatfield and Erika Yates'
date: "February 9, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 9, digits = 5)

```

These studies evaluate the effects of knockdown of the dTORC1 negative regulator *Tsc1*.  This data is located in `r getwd()` and was most recently updated on `r date()`.

## Experimental Design

```{r data-input, echo=FALSE}
require(lubridate, quietly=T)

mapping_data <- read.csv('../Data/gene_mapping.csv')
#this contains the data with the control flies
lifespan_data_file_name <- '../Data/Death Worksheet Summer 2014.csv'
lifespan.data <- read.csv(lifespan_data_file_name)
#get data from what crosses are what
lifespan_data_file_name_births <- '../Data/mhc_birth_worksheet.csv'
lifespan.data.births <- read.csv(lifespan_data_file_name_births )
lifespan.data.births$Total <- rowSums(lifespan.data.births[,10:14], na.rm=T)
all.flies <- sum(lifespan.data.births$Total, na.rm=T)

lifespan.data.births$Cross <- as.factor(paste(lifespan.data.births$Male, lifespan.data.births$Female, sep=" x "))
lifespan.data.births$Driver <- as.factor(sapply(strsplit(as.character(lifespan.data.births$Female), "/"),"[", 1))
#releveled UAS showing TSC1 then Raptor
lifespan.data.births$UAS <- as.factor(sapply(strsplit(as.character(lifespan.data.births$Male), "/"),"[", 1))
#remove analyses where incomplete data is presented
lifespan.data.births.complete <- lifespan.data.births[!(is.na(lifespan.data.births$Total)),]
#merged with gene mapping file
annotated.data <- merge(mapping_data, lifespan.data.births.complete, by="UAS", all.y=T)
annotations <- annotated.data[,c('UAS','Gene','Driver','CrossName')]
annotations <- unique(annotations[c('UAS','Gene','Driver','CrossName')])

annotated.lifespan.data <- merge(lifespan.data, annotations, by='CrossName')
#set UAS as a factor
annotated.lifespan.data$UAS <- as.factor(annotated.lifespan.data$UAS)
annotated.lifespan.data$Birth.Date <- mdy(annotated.lifespan.data$Birth.Date)
annotated.lifespan.data$Death.Date <- mdy(annotated.lifespan.data$Death.Date)
annotated.lifespan.data$age <- annotated.lifespan.data$Death.Date - annotated.lifespan.data$Birth.Date
annotated.lifespan.data <- droplevels(annotated.lifespan.data)
annotated.lifespan.data$UAS <- relevel(annotated.lifespan.data$UAS, ref='36304')


#extract only the data that corresponds to the control flies
control.data <- subset(annotated.lifespan.data, Gene=="Control")
#reformat age column
control.data$Age <- as.integer(control.data$age)
#changed 24B-GAL4 to 24B-Gal4
library(plyr)
control.data$Driver <- revalue(control.data$Driver , c("24B-GAL4"="24B-Gal4"))
control.data$Driver <- revalue(control.data$Driver , c("Hand-GAL4"="Hand-Gal4"))

#switched gender to male/female
control.data$Gender <- revalue(control.data$Gender, c("M"="male", "F"="female"))
detach("package:plyr", unload=TRUE)

tsc_filename <- "../Data/Longevity Analysis Data.csv"
tsc.data <- read.csv(tsc_filename)

library(dplyr)
#only pull in the knockdown flies (we are using the controls as the comparison group)
combined.data <- rbind(filter(tsc.data, Genotype=='shRNA/GAL4') %>% select(Gender, Gene, UAS, Driver, Age),
                       select(control.data, Gender, Gene, UAS, Driver, Age))
combined.data <- droplevels(combined.data)
combined.data$Gene <- relevel(combined.data$Gene, ref="Control")
combined.data$UAS <- relevel(as.factor(combined.data$UAS), ref="36304")
write.csv(combined.data, "../Data/dTORC1 Longevity Analysis Data.csv")
```

This script was most recently run on `r date()`.  There has been a total of **` dim(death_data)[1]`** deaths, with **` dim(death_data[death_data$Accidental==FALSE,])[1]`** of natural causes and **` dim(death_data[death_data$Phenotype!='?'&death_data$Accidental==FALSE,])[1]`** of identifiable genotypes excluding accidental deaths.  The oldest fly recorded so far was `r max(combined.data$Age, na.rm=T)` days old at time of death.

## Gene Level Analysis

```{r summary-statistics}
complete.data <- combined.data[!(is.na(combined.data$Age)),]

summarised.data.all <- 
  subset(complete.data, Driver=='24B-Gal4') %>%
  group_by(Gene, UAS) %>%
  summarise(Deaths = length(Age),
            Median.Age = median(Age),
            Age= mean(Age),
            Error = sd(Age, na.rm=T)/sqrt(length(Age)))

summarised.data.gender <- 
  subset(complete.data, Driver=='24B-Gal4') %>%
  group_by(Gene, UAS, Gender) %>%
  summarise(Deaths = length(Age),
            Median.Age = median(Age),
            Age= mean(Age),
            Error = sd(Age, na.rm=T)/sqrt(length(Age)))
```



## Number of Flies Examined

The total number of deaths for each cross and genotype, removing deaths that were accidental or not due to natural causes, the data is shown in the table below:

```{r selected-data}
library(tidyr)
kable(summarised.data.gender %>% select(Gene, Gender, Deaths) %>% spread(Gender, Deaths), caption = "Total Natural Deaths by Gender for Each Gene and Genotype", label = "tab:analysed-deaths-gene-gender")
```  

## Survival Analysis - 24B Driver

All of these are relative to the reference Genotype which is the knockdown (GAL4/shRNA).  The key packages used in this The summary statistics from this analysis are shown in the tables below:

```{r survival-analysis-24b-by-gene}
require(RColorBrewer)
genotype.colors <- brewer.pal(2, "Set2")
require(survival)
#only use 24b data
d24b.data <- droplevels(subset(complete.data, Driver=='24B-Gal4') )

#set up the dataframe
summary.coxph.24b.gene <- data.frame(row.names=levels(combined.data$Gene), 
                                n=rep(NA,length(levels(combined.data$Gene))), 
                                logtest.p=rep(NA,length(levels(combined.data$Gene))),
                                waldtest.p=rep(NA,length(levels(combined.data$Gene))),
                                sctest.p=rep(NA,length(levels(combined.data$Gene))))
#calculate statistics for each gene/control pair
for (gene in levels(d24b.data$Gene)[2]) {
    data.used <- droplevels(subset(d24b.data, Gene %in% c("Control",gene)))
    survobj <- with(data.used, Surv(Age))
    survfit <- survfit(survobj~Gene, data=data.used)
    #print(sprintf('Analysis of %s with 24B-GAL4 Driver', gene))
    #print(survfit(survobj~Genotype, data=data.used), print.rmean=TRUE)
    coxph.fit <- coxph(survobj~Gene, data=data.used)
    beta <- coef(coxph.fit)
    se <- sqrt(diag(coxph.fit$var))
    assign(paste('coxph.table.24b',gene,sep='.'),data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
    
    summary.coxph.24b.gene[gene,] <- c(
      n=coxph(survobj~Gene, data=data.used)$nevent,
      logtest.p=summary(coxph(survobj~Gene, data=data.used))$logtest[3],
      waldtest.p=summary(coxph(survobj~Gene, data=data.used))$wald[3],
      sctest.p=summary(coxph(survobj~Gene, data=data.used))$sctest[3])
    #summary.coxph.24b$padj <- p.adjust(summary.coxph.24b$logrank.p,n=3,method="BH")
    
    plot(survfit, main = sprintf('Survival of %s with 24B-GAL4 Driver', gene),
         xlab= 'Days',
         ylab= 'Percent Survival',
         yscale=100, 
         las = 1,
         lty=c(1,2,2),
         col= genotype.colors)
    legend("bottomleft", levels(data.used$Gene), col=genotype.colors, lty=c(1,2,2), bty="n")
}

par(mfrow=c(1,2))
#Separating out genders for this analysis
summary.coxph.24b.gene.gender <- data.frame(
                row.names=paste(rep(c(levels(d24b.data$Gene)[2]),2), c("male","female")),
                n=rep(NA,2), 
                logtest.p=rep(NA,2),
                waldtest.p=rep(NA,2),
                sctest.p=rep(NA,2))

#iterate over gender, then over each gene
for (gene in levels(d24b.data$Gene)[2]) {
  for (gender in levels(d24b.data$Gender)) {
        data.used <- droplevels(subset(d24b.data, Gender==gender&Gene %in% c("Control",gene)))
        survobj <- with(data.used, Surv(Age))
        survfit <- survfit(survobj~Gene, data=data.used)
        #print(survfit(survobj~Gene, data=data.used), print.rmean=TRUE)
        #print(coxph(survobj~Gene, data=data.used))
        
        coxph.fit <- coxph(survobj~Gene, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
        assign(paste(paste('coxph.table.24b',gene,sep='.'),gender, sep="."),
               data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
        
        summary.coxph.24b.gene.gender[paste(gene, gender),] <- c(
          n=coxph(survobj~Gene, data=data.used)$nevent,
          logtest.p=summary(coxph(survobj~Gene, data=data.used))$logtest[3],
          waldtest.p=summary(coxph(survobj~Gene, data=data.used))$wald[3],
          sctest.p=summary(coxph(survobj~Gene, data=data.used))$sctest[3])
        #summary.coxph.24b$padj <- p.adjust(summary.coxph.24b$logrank.p,n=3,method="BH")
        plot(survfit, main = sprintf('%s', paste(gene,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             lty=c(1,2,2,2),
             col= genotype.colors)
        legend("bottomleft", levels(data.used$Gene), col=genotype.colors, lty=c(1,2,2,2), bty="n")
}
}
```

```{r coxph-datatables-24b}
kable(summary.coxph.24b.gene, digits=c(0,9,9,9,9),caption = "Gene Level Tests for 24B-Gal4 Drivers", label = "tab:coxph-summary-24b")
kable(summary.coxph.24b.gene.gender, digits=c(0,9,9,9,9),caption = "Gene and Gender Level  Tests for 24B-Gal4 Drivers", label = "tab:coxph-summary-24b-gender")
```

### Analysis by shRNA Clone - 24B


```{r survival-analysis-24b-by-UAS}
#set up the dataframe
summary.coxph.24b.UAS <- data.frame(row.names=levels(combined.data$UAS), 
                                n=rep(NA,length(levels(combined.data$UAS))), 
                                logtest.p=rep(NA,length(levels(combined.data$UAS))),
                                waldtest.p=rep(NA,length(levels(combined.data$UAS))),
                                sctest.p=rep(NA,length(levels(combined.data$UAS))))
#calculate statistics for each UAS/control pair
for (uas in levels(d24b.data$UAS)[2:4]) {
    data.used <- droplevels(subset(d24b.data, UAS %in% c("36304",uas)))
    survobj <- with(data.used, Surv(Age))
    survfit <- survfit(survobj~UAS, data=data.used)
    #print(sprintf('Analysis of %s with 24B-GAL4 Driver', UAS))
    #print(survfit(survobj~Genotype, data=data.used), print.rmean=TRUE)
    coxph.fit <- coxph(survobj~UAS, data=data.used)
    beta <- coef(coxph.fit)
    se <- sqrt(diag(coxph.fit$var))
    assign(paste('coxph.table.24b',uas,sep='.'),data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
    
    summary.coxph.24b.UAS[uas,] <- c(
      n=coxph(survobj~UAS, data=data.used)$nevent,
      logtest.p=summary(coxph(survobj~UAS, data=data.used))$logtest[3],
      waldtest.p=summary(coxph(survobj~UAS, data=data.used))$wald[3],
      sctest.p=summary(coxph(survobj~UAS, data=data.used))$sctest[3])
    #summary.coxph.24b$padj <- p.adjust(summary.coxph.24b$logrank.p,n=3,method="BH")
    
    plot(survfit, main = sprintf('Survival of %s with 24B-GAL4 Driver', uas),
         xlab= 'Days',
         ylab= 'Percent Survival',
         yscale=100, 
         las = 1,
         lty=c(1,2,2),
         col= genotype.colors)
    legend("bottomleft", levels(data.used$UAS), col=genotype.colors, lty=c(1,2,2), bty="n")
}

par(mfrow=c(1,2))
#Separating out genders for this analysis
summary.coxph.24b.UAS.gender <- data.frame(
                row.names=as.character(outer(levels(d24b.data$Gender), levels(d24b.data$UAS),"paste")),
                n=rep(NA,8), 
                logtest.p=rep(NA,8),
                waldtest.p=rep(NA,8),
                sctest.p=rep(NA,8))

#iterate over gender, then over each UAS
for (uas in levels(d24b.data$UAS)[2:4]) {
  for (gender in levels(d24b.data$Gender)) {
        data.used <- droplevels(subset(d24b.data, Gender==gender&UAS %in% c("36304",uas)))
        survobj <- with(data.used, Surv(Age))
        survfit <- survfit(survobj~UAS, data=data.used)
        #print(survfit(survobj~UAS, data=data.used), print.rmean=TRUE)
        #print(coxph(survobj~UAS, data=data.used))
        
        coxph.fit <- coxph(survobj~UAS, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
        assign(paste(paste('coxph.table.24b',uas,sep='.'),gender, sep="."),
               data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
        
        summary.coxph.24b.UAS.gender[paste(gender, uas),] <- c(
          n=coxph(survobj~UAS, data=data.used)$nevent,
          logtest.p=summary(coxph(survobj~UAS, data=data.used))$logtest[3],
          waldtest.p=summary(coxph(survobj~UAS, data=data.used))$wald[3],
          sctest.p=summary(coxph(survobj~UAS, data=data.used))$sctest[3])
        #summary.coxph.24b$padj <- p.adjust(summary.coxph.24b$logrank.p,n=3,method="BH")
        plot(survfit, main = sprintf('%s', paste(uas,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             lty=c(1,2,2,2),
             col= genotype.colors)
        legend("bottomleft", levels(data.used$UAS), col=genotype.colors, lty=c(1,2,2,2), bty="n")
}
}
```

```{r coxph-datatables-uas-24b}
kable(summary.coxph.24b.UAS, digits=c(0,9,9,9,9),caption = "UAS Level Tests for 24B-Gal4 Drivers", label = "tab:coxph-summary-24b")
kable(summary.coxph.24b.UAS.gender, digits=c(0,9,9,9,9),caption = "UAS and Gender Level  Tests for 24B-Gal4 Drivers", label = "tab:coxph-summary-24b-gender")
```

### Hazard Ratios and Summary Tables - 24B

```{r hazard-ratios-24b}
hazard.table <- data.frame()
for (table in ls()[grep("coxph.table.24b", ls())] ){
  hazard.table <- rbind(hazard.table, cbind(group=strsplit(table, "24b.")[[1]][2], get(table)))
}
hazard.table$padj <- p.adjust(hazard.table$p, method="BH")
kable(hazard.table, caption="Hazard Ratios for All Comparasons", row.names=F)
```

## Survival Analysis - Hand

```{r survival-analysis-hand-by-gene}
require(RColorBrewer)
genotype.colors <- brewer.pal(2, "Set2")
require(survival)
#only use hand data
dhand.data <- droplevels(subset(complete.data, Driver=='Hand-Gal4') )

#set up the dataframe
summary.coxph.hand.gene <- data.frame(row.names=levels(combined.data$Gene), 
                                n=rep(NA,length(levels(combined.data$Gene))), 
                                logtest.p=rep(NA,length(levels(combined.data$Gene))),
                                waldtest.p=rep(NA,length(levels(combined.data$Gene))),
                                sctest.p=rep(NA,length(levels(combined.data$Gene))))
#calculate statistics for each gene/control pair
for (gene in levels(dhand.data$Gene)[2]) {
    data.used <- droplevels(subset(dhand.data, Gene %in% c("Control",gene)))
    survobj <- with(data.used, Surv(Age))
    survfit <- survfit(survobj~Gene, data=data.used)
    #print(sprintf('Analysis of %s with hand-GAL4 Driver', gene))
    #print(survfit(survobj~Genotype, data=data.used), print.rmean=TRUE)
    coxph.fit <- coxph(survobj~Gene, data=data.used)
    beta <- coef(coxph.fit)
    se <- sqrt(diag(coxph.fit$var))
    assign(paste('coxph.table.hand',gene,sep='.'),data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
    
    summary.coxph.hand.gene[gene,] <- c(
      n=coxph(survobj~Gene, data=data.used)$nevent,
      logtest.p=summary(coxph(survobj~Gene, data=data.used))$logtest[3],
      waldtest.p=summary(coxph(survobj~Gene, data=data.used))$wald[3],
      sctest.p=summary(coxph(survobj~Gene, data=data.used))$sctest[3])
    #summary.coxph.hand$padj <- p.adjust(summary.coxph.hand$logrank.p,n=3,method="BH")
    
    plot(survfit, main = sprintf('Survival of %s with Hand-GAL4 Driver', gene),
         xlab= 'Days',
         ylab= 'Percent Survival',
         yscale=100, 
         las = 1,
         lty=c(1,2,2),
         col= genotype.colors)
    legend("bottomleft", levels(data.used$Gene), col=genotype.colors, lty=c(1,2,2), bty="n")
}

par(mfrow=c(1,2))
#Separating out genders for this analysis
summary.coxph.hand.gene.gender <- data.frame(
                row.names=paste(rep(c(levels(dhand.data$Gene)[2]),2), c("male","female")),
                n=rep(NA,2), 
                logtest.p=rep(NA,2),
                waldtest.p=rep(NA,2),
                sctest.p=rep(NA,2))

#iterate over gender, then over each gene
for (gene in levels(dhand.data$Gene)[2]) {
  for (gender in levels(dhand.data$Gender)) {
        data.used <- droplevels(subset(dhand.data, Gender==gender&Gene %in% c("Control",gene)))
        survobj <- with(data.used, Surv(Age))
        survfit <- survfit(survobj~Gene, data=data.used)
        #print(survfit(survobj~Gene, data=data.used), print.rmean=TRUE)
        #print(coxph(survobj~Gene, data=data.used))
        
        coxph.fit <- coxph(survobj~Gene, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
        assign(paste(paste('coxph.table.hand',gene,sep='.'),gender, sep="."),
               data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
        
        summary.coxph.hand.gene.gender[paste(gene, gender),] <- c(
          n=coxph(survobj~Gene, data=data.used)$nevent,
          logtest.p=summary(coxph(survobj~Gene, data=data.used))$logtest[3],
          waldtest.p=summary(coxph(survobj~Gene, data=data.used))$wald[3],
          sctest.p=summary(coxph(survobj~Gene, data=data.used))$sctest[3])
        #summary.coxph.hand$padj <- p.adjust(summary.coxph.hand$logrank.p,n=3,method="BH")
        plot(survfit, main = sprintf('%s', paste(gene,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             lty=c(1,2,2,2),
             col= genotype.colors)
        legend("bottomleft", levels(data.used$Gene), col=genotype.colors, lty=c(1,2,2,2), bty="n")
}
}
```

```{r coxph-datatables-hand}
kable(summary.coxph.hand.gene, digits=c(0,9,9,9,9),caption = "Gene Level Tests for Hand-Gal4 Drivers", label = "tab:coxph-summary-hand")
kable(summary.coxph.hand.gene.gender, digits=c(0,9,9,9,9),caption = "Gene and Gender Level  Tests for Hand-Gal4 Drivers", label = "tab:coxph-summary-hand-gender")
```

## Analysis by shRNA Clone


```{r survival-analysis-hand-by-UAS}
#set up the dataframe
summary.coxph.hand.UAS <- data.frame(row.names=levels(combined.data$UAS), 
                                n=rep(NA,length(levels(combined.data$UAS))), 
                                logtest.p=rep(NA,length(levels(combined.data$UAS))),
                                waldtest.p=rep(NA,length(levels(combined.data$UAS))),
                                sctest.p=rep(NA,length(levels(combined.data$UAS))))
#calculate statistics for each UAS/control pair
for (uas in levels(dhand.data$UAS)[2:3]) {
    data.used <- droplevels(subset(dhand.data, UAS %in% c("36304",uas)))
    survobj <- with(data.used, Surv(Age))
    survfit <- survfit(survobj~UAS, data=data.used)
    #print(sprintf('Analysis of %s with hand-GAL4 Driver', UAS))
    #print(survfit(survobj~Genotype, data=data.used), print.rmean=TRUE)
    coxph.fit <- coxph(survobj~UAS, data=data.used)
    beta <- coef(coxph.fit)
    se <- sqrt(diag(coxph.fit$var))
    assign(paste('coxph.table.hand',uas,sep='.'),data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
    
    summary.coxph.hand.UAS[uas,] <- c(
      n=coxph(survobj~UAS, data=data.used)$nevent,
      logtest.p=summary(coxph(survobj~UAS, data=data.used))$logtest[3],
      waldtest.p=summary(coxph(survobj~UAS, data=data.used))$wald[3],
      sctest.p=summary(coxph(survobj~UAS, data=data.used))$sctest[3])
    #summary.coxph.hand$padj <- p.adjust(summary.coxph.hand$logrank.p,n=3,method="BH")
    
    plot(survfit, main = sprintf('Survival of %s with Hand-GAL4 Driver', uas),
         xlab= 'Days',
         ylab= 'Percent Survival',
         yscale=100, 
         las = 1,
         lty=c(1,2,2),
         col= genotype.colors)
    legend("bottomleft", levels(data.used$UAS), col=genotype.colors, lty=c(1,2,2), bty="n")
}

par(mfrow=c(1,2))
#Separating out genders for this analysis
summary.coxph.hand.UAS.gender <- data.frame(
                row.names=as.character(outer(levels(dhand.data$Gender), levels(dhand.data$UAS),"paste")),
                n=rep(NA,14), 
                logtest.p=rep(NA,14),
                waldtest.p=rep(NA,14),
                sctest.p=rep(NA,14))

#iterate over gender, then over each UAS
for (uas in levels(dhand.data$UAS)[2:3]) {
  for (gender in levels(dhand.data$Gender)) {
        data.used <- droplevels(subset(dhand.data, Gender==gender&UAS %in% c("36304",uas)))
        survobj <- with(data.used, Surv(Age))
        survfit <- survfit(survobj~UAS, data=data.used)
        #print(survfit(survobj~UAS, data=data.used), print.rmean=TRUE)
        #print(coxph(survobj~UAS, data=data.used))
        
        coxph.fit <- coxph(survobj~UAS, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
        assign(paste(paste('coxph.table.hand',uas,sep='.'),gender, sep="."),
               data.frame(Coef=beta, 
                              SE = se,
                              Hazard.Ratio = 1/exp(beta),
                              p = 1 - pchisq((beta/se)^2, 1)))
        
        summary.coxph.hand.UAS.gender[paste(gender, uas),] <- c(
          n=coxph(survobj~UAS, data=data.used)$nevent,
          logtest.p=summary(coxph(survobj~UAS, data=data.used))$logtest[3],
          waldtest.p=summary(coxph(survobj~UAS, data=data.used))$wald[3],
          sctest.p=summary(coxph(survobj~UAS, data=data.used))$sctest[3])
        #summary.coxph.hand$padj <- p.adjust(summary.coxph.hand$logrank.p,n=3,method="BH")
        plot(survfit, main = sprintf('%s', paste(uas,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             lty=c(1,2,2,2),
             col= genotype.colors)
        legend("bottomleft", levels(data.used$UAS), col=genotype.colors, lty=c(1,2,2,2), bty="n")
}
}
```

```{r coxph-datatables-uas-hand}
kable(summary.coxph.hand.UAS, digits=c(0,9,9,9,9),caption = "UAS Level Tests for Hand-Gal4 Drivers", label = "tab:coxph-summary-hand")
kable(summary.coxph.hand.UAS.gender, digits=c(0,9,9,9,9),caption = "UAS and Gender Level  Tests for Hand-Gal4 Drivers", label = "tab:coxph-summary-hand-gender")
```

## Hazard Ratios and Summary Tables

```{r hazard-ratios-hand}
hazard.table <- data.frame()
for (table in ls()[grep("coxph.table.hand", ls())] ){
  hazard.table <- rbind(hazard.table, cbind(group=strsplit(table, "hand.")[[1]][2], get(table)))
}
hazard.table$padj <- p.adjust(hazard.table$p, method="BH")
kable(hazard.table, caption="Hazard Ratios for All Comparasons", row.names=F)
```


## Session Information

```{r sessionInfo}
sessionInfo()
```