---
title: "Longevity Analysis of MHC-GAL4 Flies"
author: "Erika Yates and Dave Bridges"
date: "December 4, 2014"
output:
  html_document:
    keep_md: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figure/',
               echo=FALSE, warning=FALSE, message=FALSE)
```

This was based on the incomplete lifespan data from the summer of 2014.

```{r data-entry}
color.scheme <- c('black','red')

mapping.data <- read.csv('../Data/gene_mapping.csv')
mapping.data$UAS <- as.factor(mapping.data$UAS)
lifespan_data_file_name <- '../Data/Death Worksheet Summer 2014.csv'
lifespan.data <- read.csv(lifespan_data_file_name)


#get data from what crosses are what
data_file_name <- '../Data/mhc_birth_worksheet.csv'
data <- read.csv(data_file_name)
data$Total <- rowSums(data[,10:14], na.rm=T)
all.flies <- sum(data$Total, na.rm=T)

data$Cross <- as.factor(paste(data$Male, data$Female, sep=" x "))
data$Driver <- as.factor(sapply(strsplit(as.character(data$Female), "/"),"[", 1))
#releveled UAS showing TSC1 then Raptor
data$UAS <- as.factor(sapply(strsplit(as.character(data$Male), "/"),"[", 1))
#remove analyses where incomplete data is presented
data.complete <- data[!(is.na(data$Total)),]
#merged with gene mapping file
annotated.data <- merge(mapping.data, data.complete, by="UAS", all.y=T)
annotations <- annotated.data[,c('UAS','Gene','Driver','CrossName')]
annotations <- unique(annotations[c('UAS','Gene','Driver','CrossName')])

annotated.lifespan.data <- merge(lifespan.data, annotations, by='CrossName')
#set UAS as a factor
annotated.lifespan.data$UAS <- as.factor(annotated.lifespan.data$UAS)
require(lubridate)
annotated.lifespan.data$Birth.Date <- mdy(annotated.lifespan.data$Birth.Date)
annotated.lifespan.data$Death.Date <- mdy(annotated.lifespan.data$Death.Date)
annotated.lifespan.data$age <- annotated.lifespan.data$Death.Date - annotated.lifespan.data$Birth.Date
annotated.lifespan.data <- droplevels(annotated.lifespan.data)
annotated.lifespan.data$UAS <- relevel(annotated.lifespan.data$UAS, ref='36304')
```

# *Mhc*-GAL4 Crosses

## *Raptor* Knockdowns

```{r mhc-gal4-raptor-lifespan, results='asis'}
mhc_raptor_data <- droplevels(subset(annotated.lifespan.data, Driver=="Mhc-GAL4"&Gene%in%c('Control','Raptor')))

se <- function(x) sd(x)/sqrt(length(x))
library(dplyr)
mhc_raptor_summary <-
  mhc_raptor_data %>%
  group_by(Gene, Gender, UAS) %>%
  summarize(mean=mean(age),
            median = median(age),
            se = se(age),
            n = length(age))

library(xtable)
print(xtable(as.data.frame(mhc_raptor_summary), caption="Summary statistics for effects of Mhc-GAL4 > Raptor knockdown flies"), type='html')
```

We looked at a total of `r sum(mhc_raptor_summary$n)` flies in this group, with at least `r min(mhc_raptor_summary$n)` flies of each gender for each UAS.  We used the survival package to test if these changes are significant

## *Mhc*-GAL4 > *Raptor* Survival Analysis

```{r mhc-gal4-raptor-survival, results='asis', dev=c('png','pdf')}
library(survival)
par(mfrow=c(2,4))
coxph.table <- data.frame()
summary.coxph.table <- data.frame()
for (gender in levels(mhc_raptor_data$Gender)) {
    for (uas in levels(mhc_raptor_data$UAS)[2:5]) {
        data.used <- droplevels(mhc_raptor_data[mhc_raptor_data$UAS%in%c('36304',uas)&mhc_raptor_data$Gender==gender,])
        survobj <- with(data.used, Surv(as.numeric(age)))
        survfit.data <- survfit(survobj~UAS, data=data.used)

        coxph.fit <- coxph(survobj~UAS, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
coxph.table[paste(uas,gender),'Coef'] <- coef(coxph.fit)
coxph.table[paste(uas,gender),'SE'] <- sqrt(diag(coxph.fit$var))
coxph.table[paste(uas,gender),'p'] <- 1 - pchisq((beta/se)^2, 1)
coxph.table$padj <- p.adjust(coxph.table$p, method="BH")
coxph.table$Hazard <- exp(coxph.table$Coef)
        
summary.coxph.table[paste(uas, gender),'n'] <- coxph(survobj~UAS, data=data.used)$nevent
summary.coxph.table[paste(uas, gender),'logtest'] <- summary(coxph(survobj~UAS, data=data.used))$logtest[3]
summary.coxph.table[paste(uas, gender),'wald'] <- summary(coxph(survobj~UAS, data=data.used))$wald[3]
summary.coxph.table[paste(uas, gender),'sctest'] <- summary(coxph(survobj~UAS, data=data.used))$sctest[3]

        plot(survfit.data, 
             main = sprintf(paste(uas,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             col= color.scheme)
        legend("bottomleft", c("Control","Knockdown"), col=color.scheme, lty=1, bty="n")
}
}
#print(xtable(summary.coxph.table, digits=c(0,0,5,5,5)), type='html')
print(xtable(coxph.table, digits=c(3,3,3,5,5,3)), type='html')
```

## *Tsc1* Knockdowns

```{r mhc-gal4-tsc-lifespan, results='asis'}
mhc_tsc1_data <- droplevels(subset(annotated.lifespan.data, Driver=="Mhc-GAL4"&Gene%in%c('Control','Tsc1')))
se <- function(x) sd(x)/sqrt(length(x))
mhc_tsc1_summary <-
  mhc_tsc1_data %>%
  group_by(Gene, Gender, UAS) %>%
  summarize(mean=mean(age),
            median = median(age),
            se = se(age),
            n = length(age))

library(xtable)
print(xtable(as.data.frame(mhc_tsc1_summary), caption="Summary statistics for effects of Mhc-GAL4 > Tsc1 knockdown flies"), type='html')
```

We looked at a total of `r sum(mhc_tsc1_summary$n)` flies in this group, with at least `r min(mhc_tsc1_summary$n)` flies of each gender for each UAS.  We used the survival package to test if these changes are significant

## *Mhc*-GAL4 > *Tsc1* Survival Analysis

```{r mhc-gal4-tsc1-survival, results='asis', dev=c('png','pdf')}

par(mfrow=c(2,3))
coxph.table <- data.frame()
summary.coxph.table <- data.frame()
for (gender in levels(mhc_tsc1_data$Gender)) {
    for (uas in levels(mhc_tsc1_data$UAS)[2:4]) {
        data.used <- droplevels(mhc_tsc1_data[mhc_tsc1_data$UAS%in%c('36304',uas)&mhc_tsc1_data$Gender==gender,])
        survobj <- with(data.used, Surv(as.numeric(age)))
        survfit.data <- survfit(survobj~UAS, data=data.used)

        coxph.fit <- coxph(survobj~UAS, data=data.used)
        beta <- coef(coxph.fit)
        se <- sqrt(diag(coxph.fit$var))
coxph.table[paste(uas,gender),'Coef'] <- coef(coxph.fit)
coxph.table[paste(uas,gender),'SE'] <- sqrt(diag(coxph.fit$var))
coxph.table[paste(uas,gender),'p'] <- 1 - pchisq((beta/se)^2, 1)
coxph.table$padj <- p.adjust(coxph.table$p, method="BH")
coxph.table$Hazard <- exp(coxph.table$Coef)
        
summary.coxph.table[paste(uas, gender),'n'] <- coxph(survobj~UAS, data=data.used)$nevent
summary.coxph.table[paste(uas, gender),'logtest'] <- summary(coxph(survobj~UAS, data=data.used))$logtest[3]
summary.coxph.table[paste(uas, gender),'wald'] <- summary(coxph(survobj~UAS, data=data.used))$wald[3]
summary.coxph.table[paste(uas, gender),'sctest'] <- summary(coxph(survobj~UAS, data=data.used))$sctest[3]

        plot(survfit.data, 
             main = sprintf(paste(uas,gender)),
             xlab= 'Days',
             ylab= 'Percent Survival',
             yscale=100, 
             las = 1,
             col= color.scheme)
        legend("bottomleft", c("Control","Knockdown"), col=color.scheme, lty=1, bty="n")
}
}
#print(xtable(summary.coxph.table, digits=c(0,0,5,5,5)), type='html')
print(xtable(coxph.table, digits=c(3,3,3,5,5,3)), type='html')
```



## Session Information
```{r sessionInfo, echo=FALSE}
sessionInfo()
```