---
title: "Aging Analysis of Mck-TSC1"
author: "Dave Bridges, Kaleigh Fisher and Binbin Lu"
date: "February 9, 2015"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,dev=c('png','pdf'))
options(scipen = 9, digits = 5)
```

# Data Entry

These data are accumulated from the database.  The analysis includes all alive animals, animals which were sacrificed and animals which died of natural causes (denoted in the database as "Unknown").  Animals which died with an estimated death date are excluded from the analysis.  We are testing the effects of age on death by natural causes.  This script is located in `r getwd()` and was most recently run on `r date()`.
```{r data files}
data_file = "../Data/Mouse Log.csv"
# write.csv(data, data_file)
data <- read.csv(data_file)
data$Genotype <- relevel(as.factor(data$Genotype), ref="fl/fl; Tg/+")
data.useful <- subset(data, Cause_of_Death != c("Estimated"))
useful <- c("fl/fl; Tg/+", "fl/fl; +/+", "+/+; +/+", "+/+; Tg/+")
data.filtered <- subset(data.useful, Genotype %in% useful)
ko <- c("fl/fl; Tg/+", "fl/fl; +/+")
data.filtered.ko <- subset(data.useful, Genotype %in% ko)
```

# Mck-TSC1 Mice
```{r mck-subset}
data.known.mck <- subset(data.filtered, strain == "Muscle TSC1",)
data.known.mck <- droplevels(data.known.mck) # removes unused levels
data.known.mck.ko <- subset(data.filtered.ko, strain == "Muscle TSC1",)
data.known.mck.ko <- droplevels(data.known.mck.ko) # removes unused levels
```

## Analysis

The data is saved in `r getwd()` with the data saved as `r data_file` and analysed using R \cite{base}.
The data was analysed using the survival package \cite{survival1, survival2}.  Log rank tests were performed using the coin package \cite{coin1, coin2}.
This plot analyses all of the natural deaths (marked in the database as unknown)

```{r data-analysis-all}
library(survival)
data.genotyped.mck <- subset(data.known.mck, Genotype != "N.D.")
survobj.mck.all <- with(data.genotyped.mck, Surv(age, Cause_of_Death=="Unknown"))
surv.fit.mck.all <- survfit(survobj.mck.all~strain, data=data.genotyped.mck)
plot(surv.fit.mck.all)
```
This analysis contains a total of **`r surv.fit.mck.all$n`** animals, from which we have detected **`r sum(surv.fit.mck.all$n.event)`** natural deaths.  See Table below for a summary of natural deaths and see the figure below for the combined death curves with errors.

## Comparing all Four Genotypes
This analysis looks at all four genotypes for *Ckmm-Tsc1*.

```{r data-analysis}
#for all four genotypes
survobj.mck <- with(data.known.mck, Surv(age,Cause_of_Death=="Unknown"))
surv.fit.mck <- survfit(survobj.mck~Genotype, data=data.known.mck)
sdf.mck <- survdiff(formula=survobj.mck~Genotype, data=data.known.mck)
coxph.mck <- coxph(formula=survobj.mck~Genotype, data=data.known.mck)
#logrank test
library(coin)
logrank_test(formula=survobj.mck~Genotype, data=data.known.mck)
```

The chi-squared test for comparing all four genotypes is significant, with a p-value of `r pchisq(sdf.mck$chisq, df=3, lower=FALSE)`.  The results of these tests are in the table below.  The effects of each genotype, relative to the knockout strains are in Table \ref{tab:mck-coef}. These data are visualised in the figure.  This means that the knockout mice are `r 1/exp(max(coef(coxph.mck)))` to `r 1/exp(min(coef(coxph.mck)))` times more likely to die at any given time, depending on the strain.

```{r mck-data-table}
summary.mck.tests <- rbind(summary(coxph.mck)$logtest, summary(coxph.mck)$waldtest, summary(coxph.mck)$sctest)
rownames(summary.mck.tests) <- c("Likelihood ratio test", "Wald test", "Score (logrank) test")
kable(summary.mck.tests, digits=c(0,4,6), caption = "Muscle TSC1 Knockout Tests", label = "tab:mck-tests")

# Prepare the columns
beta <- coef( coxph.mck)
se   <- sqrt(diag(coxph.mck$var))
p    <- 1 - pchisq((beta/se)^2, 1)
CI   <- round(confint(coxph.mck), 3)

# Bind columns together, and select desired rows
res <- cbind(beta, se = exp(beta), CI, p)
# Print results in a LaTeX-ready form
kable(res, digits=c(3,3,3,3,6), caption = "Muscle TSC1 Knockout Coefficients, relative to Knockout", label = "tab:mck-coef")
```

# Comparing Floxed to Knockout
This section only compares fl/fl;+/+ to fl/fl;Tg/+.
```{r wt_ko}
#for wt v ko
survobj.mck.ko <- with(data.known.mck.ko, Surv(age,Cause_of_Death=="Unknown"))
surv.fit.mck.ko <- survfit(survobj.mck.ko~Genotype, data=data.known.mck.ko)
#summary(surv.fit.mck.ko) same as for all four genotypes
sdf.mck.ko <- survdiff(formula=survobj.mck.ko~Genotype, data=data.known.mck.ko)
coxph.mck.ko <- coxph(formula=survobj.mck.ko~Genotype, data=data.known.mck.ko)
#surv_test(formula=survobj.mck.ko~Genotype, data=data.known.mck.ko)
```

The chi-squared test for comparing the two genotypes is significant, with a p-value of `r pchisq(sdf.mck.ko$chisq, df=1, lower=FALSE)`.  The results of these tests are in the table below.  The effects of each genotype, relative to the knockout strains are in the table below. These results are presented graphically in the figure below.  This means that the knockout mice are `r 1/exp(coef(coxph.mck.ko))` times more likely to die at any given time.

```{r mck-data-table2}
library(xtable)
summary.mck.ko.tests <- rbind(summary(coxph.mck.ko)$logtest, summary(coxph.mck.ko)$waldtest, summary(coxph.mck.ko)$sctest)
rownames(summary.mck.ko.tests) <- c("Likelihood ratio test", "Wald test", "Score (logrank) test")
kable(summary.mck.ko.tests, digits=c(0,0,4), caption = "Muscle TSC1 Knockout Tests (WT vs KO)", label = "tab:mck-tests-ko")

# Prepare the columns
beta <- coef( coxph.mck.ko)
se   <- sqrt(diag(coxph.mck.ko$var))
p    <- 1 - pchisq((beta/se)^2, 1)
CI   <- round(confint(coxph.mck.ko), 3)

# Bind columns together, and select desired rows
res <- cbind(beta, se = exp(beta), CI, p)
# Print results in a LaTeX-ready form
kable(res, digits=c(3,3,3,3,4), caption = "Muscle TSC1 Knockout Coefficients, relative to Knockout", label = "tab:mck-coef")
```

# Death Logs
This table shows the age, and at risk individuals for each natural death, along with the \% survival and the confidence intervals.
```{r label=deaths}
library(ggfortify)
kable(fortify(surv.fit.mck))
```

```{r fitting-mck}
library(RColorBrewer)
colors <- brewer.pal(4, "Set1")
plot(surv.fit.mck, xlab="Survival Time in Days",
    ylab="% Surviving",
     yscale=100, 
     col=colors,
     lty = c(1,2,2,2),
     mark.time=F)
legend("bottomleft", levels(data.known.mck$Genotype), bty="n",fill=colors, lty=c(1,2,2,2))
```

```{r label=fitting-mck-ko}
plot(surv.fit.mck.ko, xlab="Survival Time in Days",
    ylab="% Surviving",
     yscale=100, 
     col=colors,
     mark.time=F)
legend("bottomleft", levels(data.known.mck.ko$Genotype), bty="n",fill=colors)
```

```{r mck-data-table3}
library(xtable)
summary.mck <- cbind(sdf.mck$n, sdf.mck$obs)
colnames(summary.mck) <- c("Total Animals", "Natural Deaths")
kable(summary.mck, caption = "Muscle TSC1 Knockout Summary", label = "tab:death-summary", digit=0)
```



## Combining the Control Mice

```{r label=fitting-mck-controls-combined}
#for wt v ko
data.known.mck$Knockout <- data.known.mck$Genotype != 'fl/fl; Tg/+'
survobj.mck.combined <- with(data.known.mck, Surv(age,Cause_of_Death=="Unknown"))
surv.fit.mck.combined <- survfit(survobj.mck.combined~Knockout, data=data.known.mck)
#summary(surv.fit.mck.combined) same as for all four genotypes
sdf.mck.combined <- survdiff(formula=survobj.mck.combined~Knockout, data=data.known.mck)
coxph.mck.combined <- coxph(formula=survobj.mck.combined~Knockout, data=data.known.mck)
#surv_test(formula=survobj.mck.combined~Genotype, data=data.known.mck)
plot(surv.fit.mck.combined, xlab="Survival Time in Days",
    ylab="% Surviving",
     yscale=100,
     col=c("black","red"),
     mark.time=F)
legend("bottomleft", c("Knockout", "Controls"), bty="n",fill=c("black","red"))
```


The chi-squared test for comparing the two genotypes is significant, with a p-value of `r pchisq(sdf.mck.combined$chisq, df=1, lower=FALSE)`.  The results of these tests are in the table below.  The effects of each genotype, relative to the knockout strains are in the table below. These results are presented graphically in the figure below.  This means that the knockout mice are `r 1/exp(coef(coxph.mck.combined))` times more likely to die at any given time.

```{r label=mck-data}
summary.mck.combined.tests <- rbind(summary(coxph.mck.combined)$logtest, summary(coxph.mck.combined)$waldtest, summary(coxph.mck.combined)$sctest)
rownames(summary.mck.combined.tests) <- c("Likelihood ratio test", "Wald test", "Score (logrank) test")
kable(summary.mck.combined.tests, digits=c(0,0,7), caption = "Muscle TSC1 Knockout Tests controls combined", label = "tab:mck-tests-combined")

# Prepare the columns
beta <- coef( coxph.mck.combined)
se   <- sqrt(diag(coxph.mck.combined$var))
p    <- 1 - pchisq((beta/se)^2, 1)
CI   <- round(confint(coxph.mck.combined), 3)

# Bind columns together, and select desired rows
res <- cbind(beta, se = exp(beta), CI, p)
# Print results in a LaTeX-ready form
kable(res, digits=c(3,3,3,3,8), caption = "Muscle TSC1 Knockout Coefficients, relative to Knockout, controls combined", label = "tab:mck-coef")
```

```{r bibliography}
require(bibtex)
write.bib(c(names(sessionInfo()$otherPkgs),'base'), file='references.bib')
```


\bibliography{references}
\bibliographystyle{unsrt}

# Session Information

```{r sessionInfo}
sessionInfo()
```
