Power Analysis for Drosophila Lifespan Experiments
====================================================

This analysis was performed by Dave Bridges on ```r date()```

The statistical power is the chance that an observation will be detected at a given false positive rate (for these calculations the false positive rate is 0.05).  The power is the same as one minus the false negative rate.  

The relative risk is the probability of an event, in this case death, relative to another group.  See http://en.wikipedia.org/wiki/Relative_risk for more details.

```{r power-analysis, echo=FALSE}
require(powerSurvEpi)

exp.group <- ctrl.group <- 500
risk.data.500 <- data.frame(risk=NA, power=NA)
for (risk in seq(1,1.99, by=0.01)) {
  power <- powerCT.default0(k=exp.group/ctrl.group, m=exp.group+ctrl.group, RR=risk, alpha = 0.05)
  risk.data.500 <- rbind(risk.data.500, c(risk,power))
}

exp.group <- ctrl.group <- 250
risk.data.250 <- data.frame(risk=NA, power=NA)
for (risk in seq(1,1.99, by=0.01)) {
 power <- powerCT.default0(k=exp.group/ctrl.group, m=exp.group+ctrl.group, RR=risk, alpha = 0.05)
 risk.data.250 <- rbind(risk.data.250, c(risk,power))
 }

exp.group <- ctrl.group <- 100
risk.data.100 <- data.frame(risk=NA, power=NA)
for (risk in seq(1,1.99, by=0.01)) {
  power <- powerCT.default0(k=exp.group/ctrl.group, m=exp.group+ctrl.group, RR=risk, alpha = 0.05)
  risk.data.100 <- rbind(risk.data.100, c(risk,power))
}

exp.group <- ctrl.group <- 50
risk.data.50 <- data.frame(risk=NA, power=NA)
for (risk in seq(1,1.99, by=0.01)) {
  power <- powerCT.default0(k=exp.group/ctrl.group, m=exp.group+ctrl.group, RR=risk, alpha = 0.05)
  risk.data.50 <- rbind(risk.data.50, c(risk,power))
}

exp.group <- ctrl.group <- 25
risk.data.25 <- data.frame(risk=NA, power=NA)
for (risk in seq(1,1.99, by=0.01)) {
  power <- powerCT.default0(k=exp.group/ctrl.group, m=exp.group+ctrl.group, RR=risk, alpha = 0.05)
  risk.data.25 <- rbind(risk.data.25, c(risk,power))
}
```
The number of samples can be determined either empirically by examining a series of sample sizes compared to the relative risk and statistical power:

```{r plots, dev=c('png','pdf'), echo=FALSE, fig.show='asis'}
library(RColorBrewer)
selected.colors <- brewer.pal(5, "Set1")
plot <- with(risk.data.500, plot(risk,power, 
                         ylab="Statistical Power", 
                         xlab="Relative Risk",
                         type="l", las=1,
                         cex.lab=1.5,
                         col=selected.colors[1]))
with(risk.data.250, lines(risk,power, col=selected.colors[2]))
with(risk.data.100, lines(risk,power, col=selected.colors[3]))
with(risk.data.50, lines(risk,power, col=selected.colors[4]))
with(risk.data.25, lines(risk,power, col=selected.colors[5]))
abline(a=0.8, b=0, lty=3)
legend("bottomright", 
       c("500 Per Genotype", "250 Per Genotype", "100 Per Genotype", "50 Per Genotype", "25 Per Genotype"),
       col=selected.colors,
       lty=1, bty="n")
```

The sample size can also be calculated for a given deisired relative risk and statistical power:

```{r sample-size}
ssizeCT.default(power=0.8, k=1, pE=1, pC=1, RR=1.5, alpha = 0.05)
```

References
-----------
```{r reference}
citation('powerSurvEpi')
citation()
```

Session Information
---------------------
```{r session-information}
sessionInfo()
```