---
title: "Quantification of Effects of Rapamycin on MyoD Levels"
author: "Dave Bridges"
date: "December 1, 2014"
output: 
  html_document:
    keep_md: true
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,fig.path='figure/')
```

C2C12 Cells were grown to confluence and then treated for 4h with 500 nM Rapamycin, 1 uM LYS6K2, or 10 uM MG132.  These data include experiments done on 2014-11-26 and 2014-11-29.  These data are located in `r getwd()` and was most recently updated on `r date()`.

```{r data-entry}
myod_data_file <- 'MyoD Quantification.txt'
myod_data_file_2 <- 'MyoD Quantification-3.txt'
myod_data <- rbind(read.csv(myod_data_file), read.csv(myod_data_file_2))

akt_data_file <- 'Akt Quantification.txt'
akt_data_file_2 <- 'Akt Quantification-2.txt'
akt_data <- rbind(read.csv(akt_data_file), read.csv(akt_data_file_2))

s6_data_file <- 'S6 Quantification.txt'
s6_data_file_2 <- 'S6 Quantification-2.txt'
s6_data <- rbind(read.csv(s6_data_file), read.csv(s6_data_file_2))
```

The input files are: `r akt_data_file` and `r myod_data_file`.

# MyoD Quantification

```{r myod-quantification, message=FALSE, warning=FALSE}
myod_data <- myod_data[grepl("Lane", myod_data$Lane.Name),]
myod_data <- myod_data[myod_data$Channel == 800,]
myod_data$Lane.Number <- as.integer(sapply(strsplit(as.character(myod_data$Lane.Name), " "),  "[", 2))
myod_data <- myod_data[!(myod_data$Lane.Name=='Lane 8'&myod_data$Scan=="2014-12-17"),]
myod_data$Protein <- rep(c('MyoD'),18)

library(reshape2)
myod_lane_summary <- dcast(myod_data, Scan + Lane.Number ~ Protein, value.var='I.I..K.Counts.')
myod_lane_summary$Treatment <- as.factor(c(rep(c("DMSO",
                                               "Rapamycin",
                                               "HS + DMSO",
                                               "HS + Rapamycin",
                                               "HS + Rapamycin + MG132",
                                               "HS + LYS6K2"),2),
                                         c("DMSO","Rapamycin",rep(c("HS + DMSO", "HS + Rapamycin"),2))))
myod_lane_summary$Treatment <- factor(myod_lane_summary$Treatment, 
                                      levels(myod_lane_summary$Treatment)[c(2,4,3,5,1,6)])

#normalized each experiment such that the HS + DMSO lane is equal to 1.
myod_lane_summary$MyoD_Norm <- c(
  myod_lane_summary[c(1:6), 'MyoD']/myod_lane_summary[myod_lane_summary$Treatment=='HS + DMSO', 'MyoD'][1], 
  myod_lane_summary[c(7:12), 'MyoD']/myod_lane_summary[myod_lane_summary$Treatment=='HS + DMSO', 'MyoD'][2],
  myod_lane_summary[c(13:18), 'MyoD']/mean(myod_lane_summary[myod_lane_summary$Treatment=='HS + DMSO', 'MyoD'][3:4]))

library(dplyr)
se <- function(x) sd(x)/sqrt(length(x))
myod_summary <-
  myod_lane_summary %>%
  group_by(Treatment) %>%
  summarize(mean=mean(MyoD_Norm),
            se = se(MyoD_Norm),
            sd = sd(MyoD_Norm),
            n = length(MyoD_Norm))
#force the HS + DMSO lane to have se = 0, since right now the se is including replicates 3 + 4
myod_summary[myod_summary$Treatment == 'HS + DMSO', 'se'] <- 0

ymax = max(myod_summary$mean + myod_summary$se)
plot <- with(myod_summary, barplot(mean, names.arg=Treatment,
                           las=1, ylim=c(0,ymax),cex.names=0.5,
                           ylab="MyoD Levels (Arbitrary Units)"))

superpose.eb <- function (x, y, ebl, ebu = ebl, length = 0.08, ...)
  arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3,
  length = length, ...)

superpose.eb(plot, myod_summary$mean, myod_summary$se)

myod_summary$Serum <- c(rep("Horse Serum",4), rep("FBS",2))
myod_summary$Drug <- c("DMSO","Rapamycin","LYS6K","Rapamycin + MG132", "DMSO", "Rapamycin")
plot <- barplot(as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='mean')[2:3]),
                        beside=T, las=1, ylab="MyoD Protein Levels")
superpose.eb(plot,
            as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='mean')[2:3]),
            as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='se')[2:3]))
library(car)
```

##Effects of Rapamycin

Based on these data, Rapamycin treatment reduced MyoD levels in the presence of Horse Serum by `r (1-myod_summary[myod_summary$Treatment=='HS + Rapamycin', 'mean'])*100`% +/- `r myod_summary[myod_summary$Treatment=='HS + Rapamycin', 'se']*100`%.  These data were normally distributed based on a Shapiro-Wilk test (p=`r shapiro.test(subset(myod_lane_summary, Treatment=="HS + Rapamycin")$MyoD_Norm)$p.value`).  The p-value by one-way *t*-test is `r t.test(subset(myod_lane_summary, Treatment=="HS + Rapamycin")$MyoD_Norm, mu=1, alternative='less')$p.value`.  

In the absence of horse serum, there was a `r myod_summary[myod_summary$Treatment=='Rapamycin', 'mean']/myod_summary[myod_summary$Treatment=='DMSO', 'mean']*100`% reduction.  These data were normally distributed, via a Shapiro-Wilk test (p=`r shapiro.test(residuals(aov(MyoD~Treatment, data=subset(myod_lane_summary, Treatment %in% c('Rapamycin','DMSO')))))$p.value`) and had equal variance based on Levene's test (p=`r leveneTest(MyoD~Treatment, data=subset(myod_lane_summary, Treatment %in% c('Rapamycin','DMSO')))$"Pr(>F)"[1]`).  The p-value from the two-way Student's *t*-test of that comparason is `r t.test(MyoD~Treatment, var.equal=T, data=subset(myod_lane_summary, Treatment %in% c('Rapamycin','DMSO')), alternative='greater')$p.value`.

##Effects of Horse Serum

The addition of horse serum caused an increase in MyoD levels of `r 1/myod_summary[myod_summary$Treatment == 'DMSO', 'mean']` fold (+/- `r myod_summary[myod_summary$Treatment == 'DMSO', 'se']` fold).  The one way *t*-test for this effect had a p-value of `r t.test(subset(myod_lane_summary, Treatment=="DMSO")$MyoD_Norm, mu=1, alternative='less')$p.value`.

# pS6/S6 Quantification

```{r s6-quantification, warning=FALSE}
#for the first experiment
s6_data <- s6_data[grepl("Lane", s6_data$Lane.Name),]
s6_data <- s6_data[s6_data$Name == "Band 1",]
s6_data$Lane.Number <- as.integer(sapply(strsplit(as.character(s6_data$Lane.Name), " "),  "[", 2))
s6_data <- s6_data[order(s6_data$Lane.Number, s6_data$Channel),]
for (row.name in rownames(s6_data)) {
  if (s6_data[row.name,'Channel'] == '700')
     s6_data[row.name,'Protein'] <- 'pS6'
  else 
    s6_data[row.name,'Protein'] <- 'S6'
}

s6_lane_summary <- dcast(s6_data, Scan + Lane.Number ~ Protein, value.var='I.I..K.Counts.', mean)
s6_lane_summary$Ratio <- s6_lane_summary$pS6/s6_lane_summary$S6 
s6_lane_summary$Treatment <- as.factor(rep(c("DMSO",
                                               "Rapamycin",
                                               "HS + DMSO",
                                               "HS + Rapamycin",
                                               "HS + Rapamycin + MG132",
                                               "HS + LYS6K2"),3))

s6_lane_summary$Treatment <- factor(s6_lane_summary$Treatment, 
                                      levels(s6_lane_summary$Treatment)[c(2,4,3,5,1,6)])
s6_lane_summary$Ratio.norm <- c(
  s6_lane_summary[c(1:6), 'Ratio']/s6_lane_summary[s6_lane_summary$Treatment=='DMSO', 'Ratio'][1], 
  s6_lane_summary[c(7:12), 'Ratio']/s6_lane_summary[s6_lane_summary$Treatment=='DMSO', 'Ratio'][2],
  s6_lane_summary[c(13:18), 'Ratio']/s6_lane_summary[s6_lane_summary$Treatment=='DMSO', 'Ratio'][3])

s6_lane_summary[is.na(s6_lane_summary$Ratio.norm),'Ratio.norm'] <- 0

s6_summary <-
  s6_lane_summary %>%
  group_by(Treatment) %>%
  summarize(mean=mean(Ratio.norm),
            se = se(Ratio.norm),
            sd = sd(Ratio.norm),
            n = length(Ratio.norm))

ymax = max(s6_summary$mean + s6_summary$se)
plot <- with(s6_summary, barplot(mean, names.arg=Treatment,
                           las=1, ylim=c(0,ymax), cex.names=0.3,
                           ylab="pS6/S6 Ratio"))

superpose.eb(plot, s6_summary$mean, s6_summary$se)
```
The addition of horse serum caused an increase in S6 phosphorylation ratio of `r s6_summary[s6_summary$Treatment == 'HS + DMSO', 'mean']` fold (+/- `r s6_summary[s6_summary$Treatment == 'HS + DMSO', 'se']` fold).  These data were normally distributed by a Shapiro-Wilk test (p=`r shapiro.test(subset(s6_lane_summary, Treatment=="HS + DMSO")$Ratio.norm)$p.value`).  The one way *t*-test for this effect had a p-value of `r t.test(subset(s6_lane_summary, Treatment=="HS + DMSO")$Ratio.norm, mu=1, alternative='less')$p.value`.


# pAkt/Akt Quantification

```{r akt-quantification, warning=FALSE}
akt_data <- akt_data[grepl("Lane", akt_data$Lane.Name),]
akt_data$Lane.Number <- as.integer(sapply(strsplit(as.character(akt_data$Lane.Name), " "),  "[", 2))

for (row.name in rownames(akt_data)) {
  if (akt_data[row.name,'Channel'] == '700')
     akt_data[row.name,'Protein'] <- 'pAkt'
  else 
    akt_data[row.name,'Protein'] <- 'Akt'
}
akt_data <- akt_data[akt_data$Lane.Name!='Lane 8',]

akt_lane_summary <- dcast(akt_data, Scan + Lane.Number ~ Protein, value.var='I.I..K.Counts.')
akt_lane_summary$Ratio <- akt_lane_summary$pAkt/akt_lane_summary$Akt 
akt_lane_summary$Treatment <- as.factor(rep(c("DMSO",
                                               "Rapamycin",
                                               "HS + DMSO",
                                               "HS + Rapamycin",
                                               "HS + Rapamycin + MG132",
                                               "HS + LYS6K2"),3))
akt_lane_summary$Treatment <- factor(akt_lane_summary$Treatment, 
                                      levels(akt_lane_summary$Treatment)[c(2,4,3,5,1,6)])
akt_lane_summary$Ratio.norm <- c(
  akt_lane_summary[c(1:6), 'Ratio']/akt_lane_summary[akt_lane_summary$Treatment=='HS + DMSO', 'Ratio'][1], 
  akt_lane_summary[c(7:12), 'Ratio']/akt_lane_summary[akt_lane_summary$Treatment=='HS + DMSO', 'Ratio'][2],
  akt_lane_summary[c(13:18), 'Ratio']/akt_lane_summary[akt_lane_summary$Treatment=='HS + DMSO', 'Ratio'][3])

akt_summary <-
  akt_lane_summary %>%
  group_by(Treatment) %>%
  summarize(mean=mean(Ratio.norm),
            se = se(Ratio.norm),
            sd = sd(Ratio.norm),
            n = length(Ratio.norm))

ymax = max(akt_summary$mean + akt_summary$se)
plot <- with(akt_summary, barplot(mean, names.arg=Treatment,
                           las=1, ylim=c(0,ymax), cex.names=0.3,
                           ylab="pAkt/Akt Ratio"))

superpose.eb(plot, akt_summary$mean, akt_summary$se)
```

The addition of horse serum caused an decrease phosphorylation ratio by `r 1-1/akt_summary[akt_summary$Treatment == 'DMSO', 'mean']` (+/- `r akt_summary[akt_summary$Treatment == 'DMSO', 'se']` fold).  These data were normally distributed based on a Shapiro-Wilk test (p=`r shapiro.test(subset(s6_lane_summary, Treatment=="HS + DMSO")$Ratio.norm)$p.value`).  The one way *t*-test for this effect had a p-value of `r t.test(subset(s6_lane_summary, Treatment=="HS + DMSO")$Ratio.norm, mu=1)$p.value`.

## General Effects of Horse Serum

```{r horse-serum, message=FALSE}
plot <- barplot(as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'mean'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'mean',][c(2:1),])),
              beside=TRUE, las=1, ylim=c(0,13),
  names.arg=c('S6','Akt'), ylab='Relative Phosphorylation')
legend("topright", c('10% FBS','2% Horse Serum'), bty='n', fill=grey.colors(2))
superpose.eb(plot,
             as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'mean'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'mean',][c(2:1),])),
             as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'se'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'se',][c(2:1),])))
```

## Combined Figure

```{r combined-figure, dev=c('png','pdf')}
par(mfrow=c(1,2))
plot <- barplot(as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='mean')[2:3]),
                        beside=T, las=1, ylab="Relative Protein Levels", main="MyoD")
legend("topleft", c('DMSO','Rapamycin'), bty='n', fill=grey.colors(2))
superpose.eb(plot,
            as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='mean')[2:3]),
            as.matrix(dcast(subset(myod_summary, Drug %in% c("DMSO","Rapamycin")), 
                        Drug~Serum, value.var='se')[2:3]))

plot <- barplot(as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'mean'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'mean',][c(2:1),])),
              beside=TRUE, las=1, ylim=c(0,13),
  names.arg=c('S6','Akt'), ylab='Relative Phosphorylation')
legend("topright", c('FBS','HS'), bty='n', fill=grey.colors(2))
superpose.eb(plot,
             as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'mean'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'mean',][c(2:1),])),
             as.matrix(cbind(
  s6_summary[grepl("DMSO", s6_summary$Treatment),'se'][c(2:1),],
  akt_summary[grepl("DMSO", akt_summary$Treatment),'se',][c(2:1),])))

```

## Session Information
```{r sessionInfo}
sessionInfo()
```
