Crawling Assays
=====================

```{r data-input, echo=FALSE}
mapping_file <- '../Data/gene_mapping.csv'
data_file_name <- '../Data/Crawling Data.csv'
birth_file_name <- '../Data/Birth Data.csv'
birth_data <- read.csv(birth_file_name)
crawling_data <- read.csv(data_file_name)
mapping.data <- read.csv(mapping_file)

manual_eclosure_crawling_file_name <- '../Data/Crawling Data - Manual Eclosure.csv'
manual_crawling_data <- read.csv(manual_eclosure_crawling_file_name)
```

The raw data saved in `r data_file_name`.  This script was most recently run on `r date()`.

Analysis
---------

```{r data-analysis, echo=FALSE, message=FALSE}
library(reshape2)
combined_crawling_data <- merge(crawling_data, dcast(birth_data, Female+Male+CrossName~Gender, value.var='Count', sum), by='CrossName', all.x=T)
combined_crawling_data$UAS <- as.factor(sapply(strsplit(as.character(combined_crawling_data$Male), '/'),"[", 1))
combined_crawling_data$Driver <- as.factor(sapply(strsplit(as.character(combined_crawling_data$Female), '/'),"[", 1))

annotated_crawling_data <- merge(mapping.data, combined_crawling_data, by='UAS', all.y=T)
annotated_crawling_data <- droplevels(annotated_crawling_data)
annotated_crawling_data$UAS <- as.factor(annotated_crawling_data$UAS)
annotated_crawling_data$UAS <- relevel(annotated_crawling_data$UAS, ref='36304')
annotated_crawling_data$Time <- as.numeric(as.character(annotated_crawling_data$Time))

#set time limit at 10s
time.limit <- 10
annotated_crawling_data$Over.Limit <- annotated_crawling_data$Time>time.limit

library(lubridate)
annotated_crawling_data$Born <- mdy(as.character(annotated_crawling_data$Born))
annotated_crawling_data$Assay.Date <- mdy(as.character(annotated_crawling_data$Assay.Date))
annotated_crawling_data$Age <- as.integer(annotated_crawling_data$Assay.Date-annotated_crawling_data$Born)
annotated_crawling_data$Age.Range <- cut(annotated_crawling_data$Age, breaks=c(min(annotated_crawling_data$Age),19, 45,max(annotated_crawling_data$Age)),dig.lab=2)
annotated_crawling_data$Assay.Group <- as.factor(annotated_crawling_data$Assay.Date)

crawling.summary.table <- data.frame(row.names=levels(annotated_crawling_data$UAS))
crawling.summary.table <- cbind(crawling.summary.table, 
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))))   
colnames(crawling.summary.table) <- levels(annotated_crawling_data$Age.Range)

crawling.summary.n <- data.frame(row.names=levels(annotated_crawling_data$UAS))
crawling.summary.n <- cbind(crawling.summary.n, 
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))))  

colnames(crawling.summary.n) <- levels(annotated_crawling_data$Age.Range)

#this generates a table with the percent of flies that do not crawl the limit and another table with all the flies in each age range
for (uas in rownames(crawling.summary.table)) {
    summary.table <- with(subset(annotated_crawling_data, Driver=='C179-Gal4'&UAS==uas), table(Over.Limit,Age.Range))
    crawling.summary.table[uas,] <- summary.table['TRUE',]/colSums(summary.table)*100
    crawling.summary.n[uas,] <- colSums(summary.table)
}

#calculates teh standard error for the counts
crawling.summary.error <- 100*sqrt(((1-crawling.summary.table/100)*crawling.summary.table/100)/crawling.summary.n)
#this stable has the left three columns as the number of f lies tested and the right hand colums being the number of flies that didnt crawl that high
crawling.counts.summary <- cbind(crawling.summary.n,crawling.summary.table*crawling.summary.n/100)
#crawling.counts.summary <- crawling.counts.summary[,c(1,3,2,4)]


crawling.fisher.tests <- data.frame(row.names=levels(annotated_crawling_data$UAS))
crawling.fisher.tests <- cbind(crawling.fisher.tests, 
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))),
                                rep(NA, length(levels(annotated_crawling_data$UAS))))   
colnames(crawling.fisher.tests) <- levels(annotated_crawling_data$Age.Range)

for (uas in rownames(crawling.counts.summary)[1:5]){
  crawling.fisher.tests[uas,1] <- fisher.test(crawling.counts.summary[c('36304',uas),1:2])$p.value
  crawling.fisher.tests[uas,2] <- fisher.test(crawling.counts.summary[c('36304',uas),3:4])$p.value
  crawling.fisher.tests[uas,3] <- fisher.test(crawling.counts.summary[c('36304',uas),5:6])$p.value
}

annotated.crawling.summary.table <- droplevels(merge(crawling.summary.table, mapping.data, by.x=0, by.y='UAS'))
annotated.crawling.summary.error <- droplevels(merge(crawling.summary.error, mapping.data, by.x=0, by.y='UAS'))
colnames(annotated.crawling.summary.table)[1] <- 'UAS'
colnames(annotated.crawling.summary.error)[1] <- 'UAS'

#reorder columns and re-arrange rows for summary.table
annotated.crawling.summary.table <- annotated.crawling.summary.table[,c('Gene','UAS', levels(annotated_crawling_data$Age.Range))]
annotated.crawling.summary.table <- annotated.crawling.summary.table[order(annotated.crawling.summary.table$Gene),]

annotated.crawling.summary.error <- annotated.crawling.summary.error[,c('Gene','UAS', levels(annotated_crawling_data$Age.Range))]
annotated.crawling.summary.error <- annotated.crawling.summary.error[order(annotated.crawling.summary.error$Gene),]

uas.anova <- aov(log(Time)~Age.Range+UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'))
uas.anova.date <- aov(Time~Assay.Date+Age.Range+UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'))
uas.anova.int <- aov(Time~Age.Range*UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'))

library(plyr)
uas.summary<- ddply(subset(annotated_crawling_data, Driver == 'C179-Gal4'), ~Age.Range+Gene+UAS, summarize,
                        mean = mean(Time, na.rm=T),
                        se = sd(Time, na.rm=T)/sqrt(length(Time)),
                        sd = sd(Time, na.rm=T),
                        rel.error = sd(Time, na.rm=T)/mean(Time, na.rm=T)*100,
                        n = length(Time))

uas.summary.gender<- ddply(subset(annotated_crawling_data, Driver == 'C179-Gal4'), ~Age.Range+Gene+UAS+Gender, summarize,
                        mean = mean(Time, na.rm=T),
                        se = sd(Time, na.rm=T)/sqrt(length(Time)),
                        sd = sd(Time, na.rm=T),
                        rel.error = sd(Time, na.rm=T)/mean(Time, na.rm=T)*100,
                        n = length(Time))

uas.summary.assay.group <- ddply(annotated_crawling_data, ~Assay.Group+Age.Range+Driver+Gene+UAS, summarize,
                        mean = mean(Time, na.rm=T),
                        se = sd(Time, na.rm=T)/sqrt(length(Time)),
                        sd = sd(Time, na.rm=T),
                        rel.error = sd(Time, na.rm=T)/mean(Time, na.rm=T)*100,
                        n = length(Time))

#did dunnet test vs the control
library(multcomp)
uas.dunnett <- glht(uas.anova.date, linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
uas.dunnett.summary <- cbind(summary(uas.dunnett)$test$coefficients, summary(uas.dunnett)$test$pvalues)
colnames(uas.dunnett.summary) <- c("Effect Size (s)", "p-value")
```


```{r timing-density-plot,dev=c('png','pdf'),echo=FALSE}
plot(density(annotated_crawling_data$Age), main="Crawling Assay Ages", las=1, xlab="Age (days)")
```


```{r barplot,dev=c('png','pdf'),echo=FALSE}
superpose.eb <- 
 function (x, y, ebl, ebu = ebl, length = 0.08, ...) 
 arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3, 
 length = length, ...)

par(mfrow=c(1,length(levels(uas.summary$Age.Range))))
for (range in levels(uas.summary$Age.Range)){
data.used <- subset(uas.summary,Age.Range==range)
par( mar=c(8,4.1,4.1,2.1))
ymax <- max(data.used$mean+data.used$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
}
```

```{r barplot-raptor,dev=c('png','pdf'),echo=FALSE}
par(mfrow=c(1,length(levels(uas.summary$Age.Range))))
for (range in levels(uas.summary$Age.Range)){
data.used <- droplevels(subset(uas.summary,Age.Range==range&Gene%in%c('Control','Raptor')))
par( mar=c(10,4.1,4.1,2.1))
ymax <- max(data.used$mean+data.used$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
}
```

```{r barplot-tsc, dev=c('png','pdf'),echo=FALSE}
par(mfrow=c(1,length(levels(uas.summary$Age.Range))))
for (range in levels(uas.summary$Age.Range)){
data.used <- droplevels(subset(uas.summary,Age.Range==range&Gene%in%c('Control','Tsc1')))
par( mar=c(10,4.1,4.1,2.1))
ymax <- max(data.used$mean+data.used$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
}
```

```{r barplot-tsc-3w, dev=c('png','pdf'),echo=FALSE}
data.used <- droplevels(subset(uas.summary,Age.Range=='(19,45]'&Gene%in%c('Control','Tsc1')))
par( mar=c(10,4.1,4.1,2.1))
ymax <- max(data.used$mean+data.used$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
```

```{r barplot-gender,dev=c('png','pdf'),echo=FALSE}

par(mfrow=c(2,length(levels(uas.summary.gender$Age.Range))))
for (range in levels(uas.summary.gender$Age.Range)){
data.used <- subset(uas.summary.gender,Age.Range==range&Gender=='Male')
par( mar=c(8,4.1,4.1,2.1))
ymax <- max(subset(uas.summary.gender,Gender=='Male')$mean+subset(uas.summary.gender,Gender=='Male')$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s, Males",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
}

for (range in levels(uas.summary.gender$Age.Range)){
data.used <- subset(uas.summary.gender,Age.Range==range&Gender=='Female')
par( mar=c(8,4.1,4.1,2.1))
ymax <- max(subset(uas.summary.gender,Gender=='Female')$mean+subset(uas.summary.gender,Gender=='Female')$se, na.rm=T)
plot <- barplot(data.used $mean,
        beside=T, las=2,
        ylim=c(0,ymax),
        names.arg=data.used$UAS,
        col=data.used$Gene,
        ylab="Time to Crawl 4cm (s)",
        xlab="shRNA Clone",
        main=sprintf("Age Range- %s, Females",range))

legend("topleft", levels(data.used$Gene), fill= palette()[1:length(levels(data.used$Gene))], bty='n')

superpose.eb(plot, 
             data.used$mean,
             data.used$se)
}
```

```{r barplot-time-limit,fig=TRUE,echo=FALSE}
max.ylim <- max(annotated.crawling.summary.table[,3:5], na.rm=T)+max(annotated.crawling.summary.error[,3:5], na.rm=T)
plot <- barplot(as.matrix(annotated.crawling.summary.table[,3:5]),
        beside=T, las=1,
        names.arg=levels(annotated_crawling_data$Age.Range),
        col=annotated.crawling.summary.table$Gene,
        ylab="Percent of Flies",
        ylim=c(0,max.ylim),
        main=sprintf("Flies Which Did Not Crawl 4cm in %s Seconds", time.limit),
        xlab="Fly Age")

legend("topleft", levels(annotated.crawling.summary.table$Gene), fill= palette()[1:length(levels(annotated.crawling.summary.table$Gene))], bty='n')

superpose.eb(plot, 
             as.matrix(annotated.crawling.summary.table[,3:5]),
             as.matrix(annotated.crawling.summary.error[,3:5]),)

```
```{r crawling-summary-tables,echo=FALSE,results='asis', message=FALSE}
require(xtable)
print(xtable(uas.summary, caption="Summary of Crawling Times per Cross", label="tab:summary"), include.rownames=F, type='html')
#print(xtable(uas.summary.assay.group, caption="Summary of Crawling Times per Cross Separated by Assay Group", label="tab:summary-assay-group"), include.rownames=F, type='html')
```

Statistical Analysis
----------------------

To test this effect, first we tested a model in which first the age range was taken into account, and then the uas was included as an interaction.  We tested the assumptions of this model by Shapiro-Wilk (p=`r shapiro.test(residuals(uas.anova))$p.value`) and found it did not have normally distributed residuals.  We therefore split the data into the three age bins and looked at them individually.

```{r statistics-grouped, echo=FALSE}
uas.aov.group.1 <- aov(Time~UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'&Age.Range==levels(annotated_crawling_data$Age.Range)[1]))
uas.aov.group.2 <- aov(Time~UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'&Age.Range==levels(annotated_crawling_data$Age.Range)[2]))
uas.aov.group.3 <- aov(Time~UAS, data=subset(annotated_crawling_data, Driver == 'C179-Gal4'&Age.Range==levels(annotated_crawling_data$Age.Range)[3]))
```

We then checked whether these had normally distributed residuals and they did not (p=`r shapiro.test(residuals(uas.aov.group.1))$p.value`, `r shapiro.test(residuals(uas.aov.group.2))$p.value` and `r shapiro.test(residuals(uas.aov.group.3))$p.value`).  We therefore did individual Wilcoxon-Rank Sum Tests comparing each genotype to the control group.

```{r wilcoxon-tests, echo=FALSE, message=FALSE}
wilcoxon.tables <- data.frame(row.names=levels(annotated_crawling_data$UAS))
for (test.group in levels(annotated_crawling_data$Age.Range)){
for (test.uas in levels(annotated_crawling_data$UAS)) {
  try(wilcoxon.tables[test.uas,test.group] <- wilcox.test(
    annotated_crawling_data[annotated_crawling_data$UAS=='36304'&annotated_crawling_data$Driver=='C179-Gal4'&annotated_crawling_data$Age.Range==test.group,]$Time,
    annotated_crawling_data[annotated_crawling_data$UAS==test.uas&annotated_crawling_data$Driver=='C179-Gal4'&annotated_crawling_data$Age.Range==test.group,]$Time)$p.value, silent=T)
}
}

raptor.wilcoxon.tables <- wilcoxon.tables[rownames(wilcoxon.tables)%in%mapping.data[mapping.data$Gene=='Raptor','UAS'],]
raptor.wilcoxon.tables$UAS <- rownames(raptor.wilcoxon.tables)
melted.raptor.wilcoxon.tables <- melt(raptor.wilcoxon.tables, value.name='pval', variable.name='age group')
melted.raptor.wilcoxon.tables$padj <- p.adjust(melted.raptor.wilcoxon.tables$pval, method="BH")
```  

```{r crawling-wilcoxon-tables,echo=FALSE,results='asis', message=FALSE}
print(xtable(wilcoxon.tables, caption="Wilcoxon Rank-Sum Tests for each group compared to the control, not adjusted for multiple observations", label='tab:wilcoxon-tests', digits=5), include.rownames=T, type='html')
print(xtable(melted.raptor.wilcoxon.tables, caption="Raptor Knockdown Wilcoxon Rank-Sum tests for each UAS compared to the control, adjusted by the method of Benjamini and Hochberg", label='tab:raptor-wilcoxon-tests', digits=5), include.rownames=T, type='html')
```

```{r crawling-fisher-tables,echo=FALSE,results='asis', message=FALSE}
print(xtable(crawling.fisher.tests, caption=sprintf("Fisher's Exact Test For Flies Which Do Not Crawl 4cm in %s Seconds", time.limit), label='tab:fisher-test-summary'), type='html')
```

Manual Eclosure Climbing Assay
------------------------------

Flies were assisted in eclosure at 10 days and then 4 days later a climbing assay was performed.


```{r manual-eclosure-climbing, echo=FALSE, dev=c('png','pdf')}
manual_climbing_summary <- ddply(manual_crawling_data, ~Genotype, summarise,
                                 mean = mean(Time),
                                 sd = sd(Time),
                                 se = sd(Time)/sqrt(length(Time)),
                                 n = length(Time),
                                 Shapiro = shapiro.test(Time)$p.value)
ymax = max(manual_climbing_summary$mean + manual_climbing_summary$se)
plot <- with(manual_climbing_summary, barplot(mean, ylab="Time to Climb 4 cm (s)", names.arg=Genotype,
                                              las=1, ylim=c(0,ymax)))
superpose.eb(plot, manual_climbing_summary$mean, manual_climbing_summary$se)
```

Since the data was not normal based on a Shapiro-Wilk test we did a Wilcoxon Rank Sum test comparing the two groups.  The p-value from this test was `r wilcox.test(Time~Genotype, data=manual_crawling_data)$p.value`.


Session Information
----------------------

```{r sessionInfo, echo=FALSE}
sessionInfo()
```