# Analysis of Birth Ratios for Muscle Tsc1/Raptor Flies -- Second Crosses

## Experimental Design

This was for the second batch of knockdown crosses in which only homozygous drivers were used.

```{r data-input, echo=FALSE}
births_worksheet_location <- 'https://docs.google.com/spreadsheet/pub?key=0Aitfmor6rCjYdHpHTk1BZVh0MFEtNDV6QVI2V2RWSEE&single=true&gid=0&output=csv' 
pupae_worksheet_location <- 'https://docs.google.com/spreadsheet/pub?key=0Aitfmor6rCjYdDVMbEVRd2NLbEpTQUM3REpKSEY3ckE&single=true&gid=0&output=csv'
mapping_file <- '../Data/gene_mapping.csv'
mapping.data <- read.csv(mapping_file)

birth_file_name <- '../Data/Birth Data.csv'
download.file(births_worksheet_location, birth_file_name, method="curl")
birth_data <- read.csv(birth_file_name)

pupae_filename <- '../Data/Pupae Inspection Data.csv'
download.file(pupae_worksheet_location, pupae_filename, method="curl")
pupae_data <- read.csv(pupae_filename)
```

These data are the raw data saved in `r birth_file_name` for the births and `r pupae_filename` for the pupae inspection.  This script was most recently run on `r date()`.

## C179-GAL4 Driver

This analysis is just for the C179 Driver.

```{r data-collection-c179, echo=FALSE}
library(reshape2)
combined_pupae_data <- merge(pupae_data, dcast(droplevels(subset(birth_data, Female != 'Mef-Gal4')), Female+Male+CrossName~Gender, value.var='Count', sum), by='CrossName', all=T)
combined_pupae_data$Live.Flies <- combined_pupae_data$Females + combined_pupae_data$Males
combined_pupae_data$Total.Pupae <- combined_pupae_data$Dead.Flies + combined_pupae_data$Empty.Larvae

#sets a sum of zero to be NA if the total number of alive flies was zero
combined_pupae_data[combined_pupae_data$Live.Flies==0&!(is.na(combined_pupae_data$Live.Flies)),]$Live.Flies <- NA
combined_pupae_data[is.na(combined_pupae_data$Live.Flies),]$Females <- NA
combined_pupae_data[is.na(combined_pupae_data$Live.Flies),]$Males <- NA

combined_pupae_data$Dead.Pupae.pct <- combined_pupae_data$Dead.Flies/combined_pupae_data$Total.Pupae*100
combined_pupae_data$UAS <- as.factor(sapply(strsplit(as.character(combined_pupae_data$Male), '/'),"[", 1))
combined_pupae_data$Driver <- as.factor(sapply(strsplit(as.character(combined_pupae_data$Female), '/'),"[", 1))

annotated_pupae_data <- merge(mapping.data, combined_pupae_data, by='UAS', all.y=T)
annotated_pupae_data <- droplevels(annotated_pupae_data)
output_pupae_file <- '../data/processed/Homozygous Crosses Pupae Data.csv'
write.csv(annotated_pupae_data, output_pupae_file)
annotated_pupae_data$UAS <- as.factor(annotated_pupae_data$UAS)
annotated_pupae_data$UAS <- relevel(annotated_pupae_data$UAS, ref='36304')

#this calculates the summary statistics for total number of flies
library(plyr)
total_fly_summary <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Live.Flies, na.rm=T),
                        se = sd(Live.Flies, na.rm=T)/sqrt(length(Live.Flies)),
                        sd = sd(Live.Flies, na.rm=T),
                        rel.error = sd(Live.Flies, na.rm=T)/mean(Live.Flies, na.rm=T)*100,                          
                        n = length(Live.Flies))
#calculated just for males
total_fly_summary_males <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Males, na.rm=T),
                        se = sd(Males, na.rm=T)/sqrt(length(Males)),
                        sd = sd(Males, na.rm=T),
                        rel.error = sd(Males, na.rm=T)/mean(Males, na.rm=T)*100,                          
                        n = length(Males))

#calculated just for females
total_fly_summary_females <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Females, na.rm=T),
                        se = sd(Females, na.rm=T)/sqrt(length(Females)),
                        sd = sd(Females, na.rm=T),
                        rel.error = sd(Females, na.rm=T)/mean(Females, na.rm=T)*100,                          
                        n = length(Females))

#this calcualtes the summary statistics for total number of pupae
total_pupae_summary_c179 <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Total.Pupae, na.rm=T),
                        se = sd(Total.Pupae, na.rm=T)/sqrt(length(Total.Pupae)),
                        sd = sd(Total.Pupae, na.rm=T),
                        rel.error = sd(Total.Pupae, na.rm=T)/mean(Total.Pupae, na.rm=T)*100,                              
                        n = length(Total.Pupae))

total_pupae_summary_c179_raptor <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'&Gene%in%c("Control","Raptor")), ~Gene+UAS, summarize,
                        mean = mean(Total.Pupae, na.rm=T),
                        se = sd(Total.Pupae, na.rm=T)/sqrt(length(Total.Pupae)),
                        sd = sd(Total.Pupae, na.rm=T),
                        rel.error = sd(Total.Pupae, na.rm=T)/mean(Total.Pupae, na.rm=T)*100,                              
                        n = length(Total.Pupae))

#this calcualtes the summary statistics for dead pupae 
dead_pupae_summary_c179 <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Dead.Pupae.pct, na.rm=T),
                        se = sd(Dead.Pupae.pct, na.rm=T)/sqrt(length(Dead.Pupae.pct)),
                        sd = sd(Dead.Pupae.pct, na.rm=T),
                        rel.error = sd(Dead.Pupae.pct, na.rm=T)/mean(Dead.Pupae.pct, na.rm=T)*100,
                        n = length(Dead.Pupae.pct))

dead_pupae_summary_c179_raptor <- ddply(subset(annotated_pupae_data, Driver=='C179-Gal4'&Gene%in%c('Control','Raptor')), ~Gene+UAS, summarize,
                        mean = mean(Dead.Pupae.pct, na.rm=T),
                        se = sd(Dead.Pupae.pct, na.rm=T)/sqrt(length(Dead.Pupae.pct)),
                        sd = sd(Dead.Pupae.pct, na.rm=T),
                        rel.error = sd(Dead.Pupae.pct, na.rm=T)/mean(Dead.Pupae.pct, na.rm=T)*100,
                        n = length(Dead.Pupae.pct))
```

```{r c179-total-pupae-statistics, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.pupae.c179.aov <- aov(Total.Pupae~Driver*UAS, data=annotated_pupae_data)
total.pupae.c179.raptor.aov <- aov(Total.Pupae~Driver*UAS, data=subset(annotated_pupae_data, Gene%in%c('Control','Raptor')))
library(car)
```

These aggregaged and annotated data are written out to the file `r output_pupae_file`.

### Analysis of Pupae

We inspected the pupae from vials, 21 days after the cross was set up and counted how many pupae were present, and how many had dead flies in them.  These data are summarized below:

```{r pupae-summary-tables-c179,echo=FALSE,results='asis', message=FALSE}
library(xtable)
print(xtable(total_pupae_summary_c179, caption="Summary of the Number of Pupae by Driver and UAS for C179-GAL4", label="tab:total-pupae-summary"), include.rownames=F, type='html')
```

```{r c179-total-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_pupae_summary_c179$mean+total_pupae_summary_c179$se, na.rm=T)
plot <- barplot(total_pupae_summary_c179$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_pupae_summary_c179$UAS,
        col=total_pupae_summary_c179$Gene,
        ylab="Total Number of Pupae",
        xlab="shRNA Clone",
        main="Total Number of Pupae Using the C179-GAL4 Driver")

legend("top", levels(total_pupae_summary_c179$Gene)[1:2], fill= palette()[1:2], bty='n')

superpose.eb <- 
 function (x, y, ebl, ebu = ebl, length = 0.08, ...) 
 arrows(x, y + ebu, x, y - ebl, angle = 90, code = 3, 
 length = length, ...)

superpose.eb(plot, 
             total_pupae_summary_c179$mean,
             total_pupae_summary_c179$se)
```


To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(total.pupae.c179.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for t hat Levene's test was `r leveneTest(Total.Pupae~UAS, data=subset(annotated_pupae_data,Driver%in%c('C179-Gal4')))$"Pr(>F)"[1]`.  

Based on this the ANOVA assumptions were met.  Therefore the p-value for effect of genotype on total pupae is **`r summary(total.pupae.c179.aov)[[1]][["Pr(>F)"]][2]`** for the UAS.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Total.Pupae~Driver, subset(annotated_pupae_data, UAS==36304))$p.value`**.  These data indicatee that there was no significant effect on the number of pupae.

### Raptor Knockdown Only
Looking only at *Raptor* knockdown flies separate from *Tsc1* knockdown flies we see the following.

```{r pupae-summary-tables-c179-raptor,echo=FALSE,results='asis', message=FALSE}
print(xtable(total_pupae_summary_c179_raptor, caption="Summary of the Number of Pupae by Driver and UAS for C179-GAL4", label="tab:total-pupae-summary"), include.rownames=F, type='html')
```

```{r c179-raptor-total-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_pupae_summary_c179_raptor$mean+total_pupae_summary_c179_raptor$se, na.rm=T)
plot <- barplot(total_pupae_summary_c179_raptor$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_pupae_summary_c179_raptor$UAS,
        col=total_pupae_summary_c179_raptor$Gene,
        ylab="Total Number of Pupae",
        xlab="shRNA Clone",
        main="Total Number of Pupae Using the C179-GAL4 Driver")

legend("topright", levels(total_pupae_summary_c179_raptor$Gene)[1:2], fill= palette()[1:2], bty='n')

superpose.eb(plot, 
             total_pupae_summary_c179_raptor$mean,
             total_pupae_summary_c179_raptor$se)
```


To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(total.pupae.c179.raptor.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for t hat Levene's test was `r leveneTest(Total.Pupae~Driver*UAS, data=subset(annotated_pupae_data,Driver=='C179-Gal4'&Gene%in%c('Control','Raptor')))$"Pr(>F)"[1]`.  

Based on this the ANOVA assumptions were met.  Therefore the p-value for effect of genotype on total pupae is **`r summary(total.pupae.c179.raptor.aov)[[1]][["Pr(>F)"]][2]`** for the UAS.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Total.Pupae~Driver, subset(annotated_pupae_data, UAS==36304&Gene%in%c('Control','Raptor')))$p.value`**.  These data indicatee that there was no significant effect on the number of pupae.


### Analysis of Dead Pupae

We next analysed how many pupae were dead, as a percentage of the total number of pupae.

```{r c179-dead-pupae-statistics, echo=FALSE, message=FALSE, warning=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
dead.pupae.aov <- aov(Dead.Pupae.pct~Driver*UAS, data=annotated_pupae_data)
dead.pupae.c179.aov <- aov(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'))
dead.pupae.c179.kw <- kruskal.test(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'))
#did dunnet test vs the control
library(multcomp)
pupae.dunnett <- glht(dead.pupae.c179.aov, linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
pupae.dunnett.summary <- cbind(summary(pupae.dunnett)$test$coefficients, summary(pupae.dunnett)$test$pvalues)
colnames(pupae.dunnett.summary) <- c("Effect Size (%)", "p-value")

dead.pupae.wilcoxon <- data.frame(row.names=levels(annotated_pupae_data$UAS))

for (uas in levels(annotated_pupae_data$UAS)) {
  if (uas %in% c('36304','35144')){
      dead.pupae.wilcoxon[uas,'pval'] <- NA
}
else
    dead.pupae.wilcoxon[uas,'pval'] <- wilcox.test(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'&UAS%in%c('36304',uas)))$p.value
}
dead.pupae.wilcoxon$padj <- p.adjust(dead.pupae.wilcoxon$pval, method="BH")
```

A summary of the percent of dead pupae for each cross is shown below:

```{r c179-pupae-summary-tables,echo=FALSE,results='asis'}
print(xtable(dead_pupae_summary_c179, caption="Summary of Percent of Dead Pupae by Driver and UAS for C179-GAL4", label="tab:dead-pupae-summary"), include.rownames=F, type='html')
```


```{r c-179-dead-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(dead_pupae_summary_c179$mean+dead_pupae_summary_c179$se, na.rm=T)
plot <- barplot(dead_pupae_summary_c179$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=dead_pupae_summary_c179$UAS,
        col=dead_pupae_summary_c179$Gene,
        ylab="Percent of Dead Pupae",
        xlab="shRNA Clone",
        main="Percent of Dead Pupae Using the C179-GAL4 Driver")

legend("topleft", levels(dead_pupae_summary_c179$Gene), fill= palette()[1:length(levels(dead_pupae_summary_c179$Gene))], bty='n')

superpose.eb(plot, 
             dead_pupae_summary_c179$mean,
             dead_pupae_summary_c179$se)
```

To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(dead.pupae.c179.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for that Levene's test was `r leveneTest(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data,Driver=="C179-Gal4"))$"Pr(>F)"[1]`, so neither the assumptions of equal variance or normality were met for these data.

Therefore rather than an ANOVA, we did a Kruskal-Wallis test.  The p-value for that comparason is **`r dead.pupae.c179.kw$p.value`** for the UAS.  To test if the driver alone had an effect with did a Wilcoxon Rank Sum Test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r wilcox.test(Dead.Pupae.pct~Driver, subset(annotated_pupae_data, UAS=36304))$p.value`**.  We therefore excluded the +/+ progeny from the analysis and compared with the *C179-Gal4/+* progeny as a control.  

Since this ANOVA was significant, we did post-hoc Wilcoxon-Rank Sum Tests to look at each strain relative to the control.  To test for differences compared to the *C179-GAL4/+* control strain, the results of this test are shown below.

```{r c179-pupae-dunnett-tables,echo=FALSE,results='asis'}
print(xtable(dead.pupae.wilcoxon, caption="Wilcoxon Rank Sum Tests for the Percent of Dead Pupae Comparing to Control (C179-Gal4/+)", label="tab:dead-pupae-wilcoxon", digits=c(0,5,5)), type='html')
```

#### Raptor Knockdown Only

```{r c179-raptor-dead-pupae-statistics, echo=FALSE, message=FALSE, warning=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
dead.pupae.raptor.aov <- aov(Dead.Pupae.pct~Driver*UAS, data=subset(annotated_pupae_data, Gene %in% c('Control','Raptor')))
dead.pupae.c179.raptor.aov <- aov(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'&Gene %in% c('Control','Raptor')))
dead.pupae.c179.raptor.kw <- kruskal.test(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'&Gene %in% c('Control','Raptor')))
#did dunnet test vs the control
library(multcomp)
pupae.dunnett.raptor <- glht(dead.pupae.c179.raptor.aov, linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
pupae.dunnett.raptor.summary <- cbind(summary(pupae.dunnett.raptor)$test$coefficients, summary(pupae.dunnett.raptor)$test$pvalues)
colnames(pupae.dunnett.raptor.summary) <- c("Effect Size (%)", "p-value")

dead.pupae.raptor.wilcoxon <- data.frame(row.names=levels(droplevels(subset(annotated_pupae_data, Driver=='C179-Gal4'&Gene %in% c('Control','Raptor')))$UAS))

for (uas in rownames(dead.pupae.raptor.wilcoxon)) {
  if (uas %in% c('36304')){
      dead.pupae.raptor.wilcoxon[uas,'pval'] <- NA
}
else
    dead.pupae.raptor.wilcoxon[uas,'pval'] <- wilcox.test(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'&UAS%in%c('36304',uas)))$p.value
}
dead.pupae.raptor.wilcoxon$padj <- p.adjust(dead.pupae.raptor.wilcoxon$pval, method="BH")
```

A summary of the percent of dead pupae for each cross is shown below:

```{r c179-raptor-pupae-summary-tables,echo=FALSE,results='asis'}
print(xtable(dead_pupae_summary_c179_raptor, caption="Summary of Percent of Dead Pupae by Driver and UAS for C179-GAL4", label="tab:dead-pupae-summary"), include.rownames=F, type='html')
```


```{r c-179-raptor-dead-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(dead_pupae_summary_c179_raptor$mean+dead_pupae_summary_c179_raptor$se, na.rm=T)
plot <- barplot(dead_pupae_summary_c179_raptor$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=dead_pupae_summary_c179_raptor$UAS,
        col=dead_pupae_summary_c179_raptor$Gene,
        ylab="Percent of Dead Pupae",
        xlab="shRNA Clone",
        main="Percent of Dead Pupae Using the C179-GAL4 Driver")

legend("topleft", levels(dead_pupae_summary_c179_raptor$Gene)[1:2], fill= palette()[1:2], bty='n')

superpose.eb(plot, 
             dead_pupae_summary_c179_raptor$mean,
             dead_pupae_summary_c179_raptor$se)
```

To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(dead.pupae.c179.raptor.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for that Levene's test was `r leveneTest(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data,Driver=="C179-Gal4"&Gene%in%c('Control','Raptor')))$"Pr(>F)"[1]`, so neither the assumptions of equal variance or normality were met for these data.

Therefore rather than an ANOVA, we did a Kruskal-Wallis test.  The p-value for that comparason is **`r dead.pupae.c179.kw$p.value`** for the UAS.  To test if the driver alone had an effect with did a Wilcoxon Rank Sum Test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r wilcox.test(Dead.Pupae.pct~Driver, subset(annotated_pupae_data, UAS=36304))$p.value`**.  We therefore excluded the +/+ progeny from the analysis and compared with the *C179-Gal4/+* progeny as a control.  

Since this ANOVA was significant, we did post-hoc Wilcoxon-Rank Sum Tests to look at each strain relative to the *C179-GAL4/+* control strain.  The range was `r (min(dead_pupae_summary_c179_raptor$mean[2:4])-dead_pupae_summary_c179_raptor$mean[1])/dead_pupae_summary_c179_raptor$mean[1]`-`r (max(dead_pupae_summary_c179_raptor$mean[2:4])-dead_pupae_summary_c179_raptor$mean[1])/dead_pupae_summary_c179_raptor$mean[1]` times more dead pupae.  The results of these tests are shown below.

```{r c179-raptor-pupae-dunnett-tables,echo=FALSE,results='asis'}
print(xtable(dead.pupae.raptor.wilcoxon, caption="Wilcoxon Rank Sum Tests for the Percent of Dead Pupae Comparing to Control (C179-Gal4/+)", label="tab:dead-pupae-wilcoxon", digits=c(0,5,5)), type='html')
```
     
### Analysis of Total Number of Flies

```{r c179-total-fly-statistics, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.flies.aov <- aov(Live.Flies~Driver*UAS, data=annotated_pupae_data)
restricted.flies.aov <- aov(Live.Flies~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'))
total.flies.dunnett <- glht(restricted.flies.aov , linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
fly.dunnett.summary <- cbind(summary(total.flies.dunnett)$test$coefficients, summary(total.flies.dunnett)$test$pvalues)
colnames(fly.dunnett.summary) <- c("Effect Size (%)", "p-value")
```


We inspected the total number of flies eclosed up to 21 days after the cross was set up.  These data are summarized in the table and figure below:

```{r c179-total-fly-summary-tables,echo=FALSE,results='asis'}
print(xtable(total_fly_summary, caption="Summary of the Number of Flies by Driver and UAS for C179-GAL4", label="tab:total-flies-summary"), include.rownames=F, type='html')
```

```{r c179-total-fly-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_fly_summary$mean+total_fly_summary$se, na.rm=T)
plot <- barplot(total_fly_summary$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary$UAS,
        col=total_fly_summary$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Flies for the C179-GAL4 Driver")

legend("top", levels(total_fly_summary$Gene), fill= palette()[1:length(levels(total_fly_summary$Gene))], bty='n')

superpose.eb(plot, 
             total_fly_summary$mean,
             total_fly_summary$se)
```

Table \ref{tab:total-flies-summary} and in Figure \ref{fig:total-flies-barplot}.  To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Live.Flies~Driver, subset(annotated_pupae_data, UAS=36304))$p.value`**.  

Due to the significant ANOVA we performed a post-hoc Dunnett's test using the *C179/+* progeny as the controls.  An ANOVA analysis excluding the +/+ flies had a p-value of **`r summary(restricted.flies.aov)[[1]][["Pr(>F)"]][1]`** for the UAS.  The results of that test are in Table \ref{tab:total-flies-dunnett}.  This test indicates that each of the three Raptor shRNA knockdowns cause a reduction in the number of flies.

### Analysis of Number of Flies by Gender

```{r c179-total-fly-statistics-gender, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.flies.aov.males <- aov(Males~Driver*UAS, data=annotated_pupae_data)
total.flies.aov.females <- aov(Females~Driver*UAS, data=annotated_pupae_data)

restricted.flies.aov.males <- aov(Males~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'))
restricted.flies.aov.females <- aov(Females~UAS, data=subset(annotated_pupae_data, Driver=='C179-Gal4'))

total.flies.dunnett.males <- glht(restricted.flies.aov.males , linfct=mcp(UAS="Dunnett"))
total.flies.dunnett.females <- glht(restricted.flies.aov.females , linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
fly.dunnett.summary.gendered <- cbind(summary(total.flies.dunnett.males)$test$coefficients, 
                             summary(total.flies.dunnett.males)$test$pvalues,
                             summary(total.flies.dunnett.females)$test$coefficients, 
                             summary(total.flies.dunnett.females)$test$pvalues)
colnames(fly.dunnett.summary.gendered) <- c("Male Effect Size (%)", "Male p-value","Female Effect Size (%)", "Female p-value")
```

For the flies which had been born, we could separate these data based on gender.  The separated data is summarized in Tables \ref{tab:total-flies-summary-males} and \ref{tab:total-flies-summary-females}.  These data are graphed in Figure \ref{fig:total-flies-barplot-gendered}.  

For Males, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov.males)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova-males}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Males~Driver, subset(annotated_pupae_data, UAS=36304))$p.value`**. 

For Females, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov.females)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova-females}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Females~Driver, subset(annotated_pupae_data, UAS=36304))$p.value`**. 

Due to the significant ANOVA we performed a post-hoc Dunnett's test using the *C179/+* progeny as the controls.  An ANOVA analysis excluding the +/+ flies had a p-value of **`r summary(restricted.flies.aov.males)[[1]][["Pr(>F)"]][1]`** for the UAS for males and a p-value of **`r summary(restricted.flies.aov.females)[[1]][["Pr(>F)"]][1]`** for the UAS for the females.  The results of that test are in Tables \ref{tab:total-flies-dunnett-gendered}.  This test indicates that each of the three Raptor shRNA knockdowns cause a reduction in the number of male flies, but only 34814 had significantly less females.

```{r c179-total-fly-summary-gender-tables,echo=FALSE,results='asis'}
print(xtable(total_fly_summary_males, caption="Summary of the Number of Male Flies by Driver and UAS for C179-GAL4", label="tab:total-flies-summary-males"), include.rownames=F, type='html')
print(xtable(total_fly_summary_females, caption="Summary of the Number of Female Flies by Driver and UAS for C179-GAL4", label="tab:total-flies-summary-females"), include.rownames=F, type='html')
print(xtable(total.flies.aov.males, caption="ANOVA for the Number of Male Flies by Driver and UAS for C179-GAL4", label="tab:total-flies-anova-males"), type='html')
print(xtable(total.flies.aov.females, caption="ANOVA for the Number of Female Flies by Driver and UAS for C179-GAL4", label="tab:total-flies-anova-females"), type='html')
print(xtable(fly.dunnett.summary.gendered, caption="Dunnett's Test for the Number of Males and Females Comparing to Control (C179-Gal4/+)", label="tab:total-flies-dunnett-gendered", digits=c(0,1,7,1,7)), type='html')
```

```{r total-fly-barplot-gendered,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
par(mfrow=c(2,1))
ymax <- max(total_fly_summary_males$mean+total_fly_summary_males$se, na.rm=T)
plot <- barplot(total_fly_summary_males$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary_males$UAS,
        col=total_fly_summary_males$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Male Flies")

legend("top", levels(total_fly_summary_males$Gene), fill= palette()[1:length(levels(total_fly_summary_males$Gene))], bty='n')

superpose.eb(plot, 
             total_fly_summary_males$mean,
             total_fly_summary_males$se)

ymax <- max(total_fly_summary_females$mean+total_fly_summary_females$se, na.rm=T)
plot <- barplot(total_fly_summary_females$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary_females$UAS,
        col=total_fly_summary_females$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Female Flies")

superpose.eb(plot, 
             total_fly_summary_females$mean,
             total_fly_summary_females$se)
```
     
## Mef-GAL4 Driver


This analysis is just for the *Mef2-GAL4* Driver.

```{r data-collection-mef, echo=FALSE}
combined_pupae_data_mef <- merge(pupae_data, dcast(droplevels(subset(birth_data, Female != 'C179-Gal4')), Female+Male+CrossName~Gender, value.var='Count', sum), by='CrossName', all=T)
combined_pupae_data_mef$Live.Flies <- combined_pupae_data_mef$Females + combined_pupae_data_mef$Males
combined_pupae_data_mef$Total.Pupae <- combined_pupae_data_mef$Dead.Flies + combined_pupae_data_mef$Empty.Larvae

#sets a sum of zero to be NA if the total number of alive flies was zero
combined_pupae_data_mef[combined_pupae_data_mef$Live.Flies==0&!(is.na(combined_pupae_data_mef$Live.Flies)),]$Live.Flies <- NA
combined_pupae_data_mef[is.na(combined_pupae_data_mef$Live.Flies),]$Females <- NA
combined_pupae_data_mef[is.na(combined_pupae_data_mef$Live.Flies),]$Males <- NA

combined_pupae_data_mef$Dead.Pupae.pct <- combined_pupae_data_mef$Dead.Flies/combined_pupae_data_mef$Total.Pupae*100
combined_pupae_data_mef$UAS <- as.factor(sapply(strsplit(as.character(combined_pupae_data_mef$Male), '/'),"[", 1))
combined_pupae_data_mef$Driver <- as.factor(sapply(strsplit(as.character(combined_pupae_data_mef$Female), '/'),"[", 1))

annotated_pupae_data_mef <- merge(mapping.data, combined_pupae_data_mef, by='UAS', all.y=T)
annotated_pupae_data_mef <- droplevels(annotated_pupae_data_mef)
annotated_pupae_data_mef$UAS <- as.factor(annotated_pupae_data_mef$UAS)
annotated_pupae_data_mef$UAS <- relevel(annotated_pupae_data_mef$UAS, ref='36304')

#this calculates the summary statistics for total number of flies
library(plyr)
total_fly_summary_mef <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Live.Flies, na.rm=T),
                        se = sd(Live.Flies, na.rm=T)/sqrt(length(Live.Flies)),
                        sd = sd(Live.Flies, na.rm=T),
                        rel.error = sd(Live.Flies, na.rm=T)/mean(Live.Flies, na.rm=T)*100,                          
                        n = length(Live.Flies))
#calculated just for males
total_fly_summary_males_mef <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Males, na.rm=T),
                        se = sd(Males, na.rm=T)/sqrt(length(Males)),
                        sd = sd(Males, na.rm=T),
                        rel.error = sd(Males, na.rm=T)/mean(Males, na.rm=T)*100,                          
                        n = length(Males))

#calculated just for females
total_fly_summary_females_mef <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Females, na.rm=T),
                        se = sd(Females, na.rm=T)/sqrt(length(Females)),
                        sd = sd(Females, na.rm=T),
                        rel.error = sd(Females, na.rm=T)/mean(Females, na.rm=T)*100,                          
                        n = length(Females))

#this calcualtes the summary statistics for total number of pupae
total_pupae_summary_mef <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Total.Pupae, na.rm=T),
                        se = sd(Total.Pupae, na.rm=T)/sqrt(length(Total.Pupae)),
                        sd = sd(Total.Pupae, na.rm=T),
                        rel.error = sd(Total.Pupae, na.rm=T)/mean(Total.Pupae, na.rm=T)*100,                              
                        n = length(Total.Pupae))

total_pupae_summary_mef_raptor <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'&Gene%in%c('Control','Raptor')), ~Gene+UAS, summarize,
                        mean = mean(Total.Pupae, na.rm=T),
                        se = sd(Total.Pupae, na.rm=T)/sqrt(length(Total.Pupae)),
                        sd = sd(Total.Pupae, na.rm=T),
                        rel.error = sd(Total.Pupae, na.rm=T)/mean(Total.Pupae, na.rm=T)*100,                              
                        n = length(Total.Pupae))

#this calcualtes the summary statistics for dead pupae 
dead_pupae_summary_mef <- ddply(subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'), ~Gene+UAS, summarize,
                        mean = mean(Dead.Pupae.pct, na.rm=T),
                        se = sd(Dead.Pupae.pct, na.rm=T)/sqrt(length(Dead.Pupae.pct)),
                        sd = sd(Dead.Pupae.pct, na.rm=T),
                        rel.error = sd(Dead.Pupae.pct, na.rm=T)/mean(Dead.Pupae.pct, na.rm=T)*100)
```

```{r mef-total-pupae-statistics, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.pupae.mef.aov <- aov(Total.Pupae~Driver*UAS, data=annotated_pupae_data_mef)
total.pupae.mef.raptor.aov <- aov(Total.Pupae~Driver*UAS, data=subset(annotated_pupae_data_mef, Gene%in%c('Control','Raptor')))
```

## Analysis of Pupae

We inspected the pupae from vials, 21 days after the cross was set up and counted how many pupae were present, and how many had dead flies in them.  These data are summarized below:

```{r mef-pupae-summary-tables,echo=FALSE,results='asis'}
print(xtable(total_pupae_summary_mef, caption="Summary of the Number of Pupae by Driver and UAS for Mef-GAL4", label="tab:total-pupae-summary-mef"), include.rownames=F, type='html')
```


```{r mef-total-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_pupae_summary_mef$mean+total_pupae_summary_mef$se, na.rm=T)
plot <- barplot(total_pupae_summary_mef$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_pupae_summary_mef$UAS,
        col=total_pupae_summary_mef$Gene,
        ylab="Total Number of Pupae",
        xlab="shRNA Clone",
        main="Total Number of Pupae Using the Mef-GAL4 Driver")

legend("top", levels(total_pupae_summary_mef$Gene), fill= palette()[1:length(levels(total_pupae_summary_mef$Gene))], bty='n')

superpose.eb(plot, 
             total_pupae_summary_mef$mean,
             total_pupae_summary_mef$se)
```

To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(total.pupae.mef.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for that Levene's test was `r leveneTest(Total.Pupae~Driver*UAS, data=annotated_pupae_data_mef)$"Pr(>F)"[1]`.  

Based on this the ANOVA assumptions were met.  Therefore the p-value for that comparason is **`r summary(total.pupae.mef.aov)[[1]][["Pr(>F)"]][2]`** for the UAS.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *Mef/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Total.Pupae~Driver, subset(annotated_pupae_data_mef, UAS==36304))$p.value`**.  This indicated that there was no significant effect on the number of pupae.

### Raptor Knockdown Only
Looking only at *Raptor* knockdown flies separate from *Tsc1* knockdown flies we see the following.

```{r pupae-summary-tables-mef-raptor,echo=FALSE,results='asis', message=FALSE}
print(xtable(total_pupae_summary_mef_raptor, caption="Summary of the Number of Pupae by Driver and UAS for C179-GAL4", label="tab:total-pupae-summary"), include.rownames=F, type='html')
```

```{r mef-raptor-total-pupae-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_pupae_summary_mef_raptor$mean+total_pupae_summary_mef_raptor$se, na.rm=T)
plot <- barplot(total_pupae_summary_mef_raptor$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_pupae_summary_mef_raptor$UAS,
        col=total_pupae_summary_c179_raptor$Gene,
        ylab="Total Number of Pupae",
        xlab="shRNA Clone",
        main="Total Number of Pupae Using the Mef-GAL4 Driver")

legend("topleft", levels(total_pupae_summary_mef_raptor$Gene)[1:2], fill= palette()[1:2], bty='n')

superpose.eb(plot, 
             total_pupae_summary_mef_raptor$mean,
             total_pupae_summary_mef_raptor$se)
```


To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  We then checked wither the residuals were normally distributed.  The p-value for that Shapiro-Wilk test was p=`r shapiro.test(residuals(total.pupae.mef.raptor.aov))$p.value`.  We also checked for equal variance between groups.  The p-value for that Levene's test was `r leveneTest(Total.Pupae~Driver*UAS, data=subset(annotated_pupae_data_mef,Gene%in%c('Control','Raptor')))$"Pr(>F)"[1]`.  

Based on this the ANOVA assumptions were met.  Therefore the p-value for effect of genotype on total pupae is **`r summary(total.pupae.mef.raptor.aov)[[1]][["Pr(>F)"]][2]`** for the UAS.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *C179/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Total.Pupae~Driver, subset(annotated_pupae_data_mef, UAS==36304&Gene%in%c('Control','Raptor')))$p.value`**.  These data indicatee that there was no significant effect on the number of pupae.


### Analysis of Dead Pupae

```{r dead-pupae-barplot-mef,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(100)
plot <- barplot(dead_pupae_summary_mef$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=dead_pupae_summary_mef$UAS,
        col=dead_pupae_summary_mef$Gene,
        ylab="Percent of Dead Pupae",
        xlab="shRNA Clone",
        main="Percent of Dead Pupae Using the Mef-GAL4 Driver")

legend("topleft", levels(dead_pupae_summary_mef$Gene), fill= palette()[1:length(levels(dead_pupae_summary_mef$Gene))], bty='n')

superpose.eb(plot, 
             dead_pupae_summary_mef$mean,
             dead_pupae_summary_mef$se)
```

```{r mef-dead-pupae-statistics, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
pupae.aov_mef <- aov(Dead.Pupae.pct~Driver*UAS, data=annotated_pupae_data_mef)
pupae.aov.Mef_mef <- aov(Dead.Pupae.pct~UAS, data=subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'))
#did dunnet test vs the control
library(multcomp)
pupae.dunnett_mef <- glht(pupae.aov.Mef_mef, linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
pupae.dunnett.summary_mef <- cbind(summary(pupae.dunnett_mef)$test$coefficients, summary(pupae.dunnett_mef)$test$pvalues)
colnames(pupae.dunnett.summary_mef) <- c("Effect Size (%)", "p-value")
```

A summary of the percent of dead pupae for each cross is shown in Table \ref{tab:dead-pupae-summary-mef} and in Figure \ref{fig:dead-pupae-barplot}.
To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the percentage of dead pupae.  The p-value for that comparason is **`r summary(pupae.aov_mef)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:dead-pupae-anova-mef}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *Mef/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Dead.Pupae.pct~Driver, subset(annotated_pupae_data_mef, UAS=36304))$p.value`**.  We therefore excluded the +/+ progeny from the analysis and compared with the *Mef-Gal4/+* progeny as a control.  

The ANOVA for this comparason yielded a p-value for the UAS of **`r summary(pupae.aov.Mef_mef)[[1]][["Pr(>F)"]][1]`**.  The results of this ANOVA are in Table \ref{tab:dead-pupae-anova-Mef}.  Since this ANOVA was significant, we did a post-hoc test to look at each strain relative to the control.  To test for differences compared to the *Mef-GAL4/+* control strain, we did a Dunnett's test.  The range was `r (min(dead_pupae_summary_mef$mean[2:3])-dead_pupae_summary_mef$mean[1])/dead_pupae_summary_mef$mean[1]`-`r (max(dead_pupae_summary_mef$mean[2:3])-dead_pupae_summary_mef$mean[1])/dead_pupae_summary_mef$mean[1]` times more dead pupae. The results of this test are shown in Table \ref{tab:dead-pupae-dunnett-mef}.  


```{r mef-dead-pupae-summary-tables,echo=FALSE,results='asis', message=FALSE}
print(xtable(dead_pupae_summary_mef, caption="Summary of Percent of Dead Pupae by Driver and UAS for Mef-GAL4", label="tab:dead-pupae-summary-mef"), include.rownames=F, type='html')
print(xtable(pupae.aov_mef, caption="ANOVA for Percent of Dead Pupae by Driver and UAS for Mef-GAL4", label="tab:dead-pupae-anova-mef"), type='html')
print(xtable(pupae.aov.Mef_mef, caption="ANOVA for Percent of Dead Pupae by UAS for Mef-GAL4", label="tab:dead-pupae-anova-Mef"), type='html')
print(xtable(pupae.dunnett.summary_mef, caption="Dunnett's Test for the Percent of Dead Pupae Comparing to Control (Mef-Gal4/+)", label="tab:dead-pupae-dunnett-mef", digits=c(0,1,7)), type='html')
```

### Analysis of Total Number of Flies}

```{r mef-total-fly-statistics, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.flies.aov_mef <- aov(Live.Flies~Driver*UAS, data=annotated_pupae_data_mef)
restricted.flies.aov_mef <- aov(Live.Flies~UAS, data=subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'))
total.flies.dunnett_mef <- glht(restricted.flies.aov_mef , linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
fly.dunnett.summary_mef <- cbind(summary(total.flies.dunnett_mef)$test$coefficients, summary(total.flies.dunnett_mef)$test$pvalues)
colnames(fly.dunnett.summary_mef) <- c("Effect Size (%)", "p-value")
```

```{r mef-total-fly-barplot,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
ymax <- max(total_fly_summary_mef$mean+total_fly_summary_mef$se, na.rm=T)
plot <- barplot(total_fly_summary_mef$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary_mef$UAS,
        col=total_fly_summary_mef$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Flies for the Mef-GAL4 Driver")

legend("top", levels(total_fly_summary_mef$Gene), fill= palette()[1:length(levels(total_fly_summary_mef$Gene))], bty='n')

superpose.eb(plot, 
             total_fly_summary_mef$mean,
             total_fly_summary_mef$se)
```

We inspected the total number of flies eclosed up to 21 days after the cross was set up.  These data are summarized in  Table \ref{tab:total-flies-summary-mef} and in Figure \ref{fig:total-flies-barplot-mef}.  To analyse these data, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov_mef)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *Mef/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Live.Flies~Driver, subset(annotated_pupae_data_mef, UAS=36304))$p.value`**.  

Due to the significant ANOVA we performed a post-hoc Dunnett's test using the *Mef/+* progeny as the controls.  An ANOVA analysis excluding the +/+ flies had a p-value of **`r summary(restricted.flies.aov_mef)[[1]][["Pr(>F)"]][1]`** for the UAS.  The results of that test are in Table \ref{tab:total-flies-dunnett-mef}.  This test indicates that each of the three Raptor shRNA knockdowns cause a reduction in the number of flies.

```{r mef-total-fly-summary-tables,echo=FALSE,results='asis'}
print(xtable(total_fly_summary_mef, caption="Summary of the Number of Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-summary-mef"), include.rownames=F, type='html')
print(xtable(total.flies.aov_mef, caption="ANOVA for the Number of Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-anova-mef"), type='html')
print(xtable(fly.dunnett.summary_mef, caption="Dunnett's Test for the Number of Flies Comparing to Control (Mef-Gal4/+)", label="tab:total-flies-dunnett-mef", digits=c(0,1,7)), type='html')
```

### Analysis of Number of Flies by Gender

```{r mef-total-fly-statistics-gender, echo=FALSE}
#perform anova testing first for an effect of the driver, then for an effect of the shRNAs
total.flies.aov.males_mef <- aov(Males~Driver*UAS, data=annotated_pupae_data_mef)
total.flies.aov.females_mef <- aov(Females~Driver*UAS, data=annotated_pupae_data_mef)

restricted.flies.aov.males_mef <- aov(Males~UAS, data=subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'))
restricted.flies.aov.females_mef <- aov(Females~UAS, data=subset(annotated_pupae_data_mef, Driver=='Mef-Gal4'))

total.flies.dunnett.males_mef <- glht(restricted.flies.aov.males_mef , linfct=mcp(UAS="Dunnett"))
total.flies.dunnett.females_mef <- glht(restricted.flies.aov.females_mef , linfct=mcp(UAS="Dunnett"))
#made a table with the dunnet test results
fly.dunnett.summary.gendered_mef <- cbind(summary(total.flies.dunnett.males_mef)$test$coefficients, 
                             summary(total.flies.dunnett.males_mef)$test$pvalues,
                             summary(total.flies.dunnett.females_mef)$test$coefficients, 
                             summary(total.flies.dunnett.females_mef)$test$pvalues)
colnames(fly.dunnett.summary.gendered_mef) <- c("Male Effect Size (%)", "Male p-value","Female Effect Size (%)", "Female p-value")
```

For the flies which had been born, we could separate these data based on gender.  The separated data is summarized in Tables \ref{tab:total-flies-summary-males} and \ref{tab:total-flies-summary-females}.  These data are graphed in Figure \ref{fig:total-flies-barplot-gendered}.  

For Males, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov.males_mef)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova-males-mef}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *Mef/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Males~Driver, subset(annotated_pupae_data_mef, UAS=36304))$p.value`**. 

For Females, first we did an ANOVA testing for whether the UAS has an effect on the number of eclosed flies.  The p-value for that comparason is **`r summary(total.flies.aov.females_mef)[[1]][["Pr(>F)"]][2]`** for the UAS.  The results of this ANOVA are shown in Table \ref{tab:total-flies-anova-females}.  To test if the driver alone had an effect with did a Student's t-test comparing the +/+ and *Mef/+* progeny.  The p-value for the Driver having an effect is **`r t.test(Females~Driver, subset(annotated_pupae_data_mef, UAS=36304))$p.value`**. 

Due to the significant ANOVA we performed a post-hoc Dunnett's test using the *Mef/+* progeny as the controls.  An ANOVA analysis excluding the +/+ flies had a p-value of **`r summary(restricted.flies.aov.males_mef)[[1]][["Pr(>F)"]][1]`** for the UAS for males and a p-value of **`r summary(restricted.flies.aov.females_mef)[[1]][["Pr(>F)"]][1]`** for the UAS for the females.  The results of that test are in Tables \ref{tab:total-flies-dunnett-gendered}.  This test indicates that each of the three Raptor shRNA knockdowns cause a reduction in the number of male flies, but only 34814 had significantly less females.

```{r mef-total-fly-summary-gender-tables,echo=FALSE,results='asis'}
print(xtable(total_fly_summary_males_mef, caption="Summary of the Number of Male Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-summary-males-mef"), include.rownames=F, type='html')
print(xtable(total_fly_summary_females_mef, caption="Summary of the Number of Female Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-summary-females-mef"), include.rownames=F, type='html')
print(xtable(total.flies.aov.males_mef, caption="ANOVA for the Number of Male Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-anova-males-mef"), type='html')
print(xtable(total.flies.aov.females_mef, caption="ANOVA for the Number of Female Flies by Driver and UAS for Mef-GAL4", label="tab:total-flies-anova-females-mef"), type='html')
print(xtable(fly.dunnett.summary.gendered_mef, caption="Dunnett's Test for the Number of Males and Females Comparing to Control (Mef-Gal4/+)", label="tab:total-flies-dunnett-gendered-mef", digits=c(0,1,7,1,7)), type='html')
```

```{r mef-total-fly-barplot-gendered,fig=TRUE,echo=FALSE, dev=c('png','pdf')}
par(mfrow=c(2,1))
ymax <- max(total_fly_summary_males_mef$mean+total_fly_summary_males_mef$se, na.rm=T)
plot <- barplot(total_fly_summary_males_mef$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary_males_mef$UAS,
        col=total_fly_summary_males_mef$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Male Flies")

legend("top", levels(total_fly_summary_males_mef$Gene), fill= palette()[1:length(levels(total_fly_summary_males_mef$Gene))], bty='n')

superpose.eb(plot, 
             total_fly_summary_males_mef$mean,
             total_fly_summary_males_mef$se)

ymax <- max(total_fly_summary_females_mef$mean+total_fly_summary_females_mef$se, na.rm=T)
plot <- barplot(total_fly_summary_females_mef$mean,
        beside=T, las=1,
        ylim=c(0,ymax),
        names.arg=total_fly_summary_females_mef$UAS,
        col=total_fly_summary_females_mef$Gene,
        ylab="Total Number of Flies",
        xlab="shRNA Clone",
        main="Total Number of Female Flies")

superpose.eb(plot, 
             total_fly_summary_females_mef$mean,
             total_fly_summary_females_mef$se)
```

## Session Information
```{r sessionInfo, echo=F}
sessionInfo()
```